### 2021-05-23 DeepLearningAndBayesianSoftware.Rmd
initial checkin, fairly complete draft

---
title: "Deep Learning and Bayesian Modeling: Comparing research use via citation counting"
author: "Breck Baldwin"
date: "5/19/2021"
output:
 bookdown::tufte_html2: default
bibliography: citations.bib
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	comment = NA
)
```

Stan, a Bayesian modeling language, was released in 2012 to considerable fanfare. A shiny new inference algorithm, HMC with NUTS (@hoffman2014no), promised and delivered fits that could not be fit before. And then deep learning happened. The relese of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to everyone and in 2016 began garnering significant traction. One result of this was people often assumed that deep learning's successes usurped Bayesian modeling's domain. This was not in our imagination, NSF reviews came back dismissing Stan funding because all the interesting work was assumed to be happening with deep learning. Mind you, deep learning and Bayesian modeling are conceptual cousins but very different from each other and are better thought of as complementary than as antagonistic. Yet Bayesians found themselves in deep learning's shadow somehow. It is 2021 and this document does some very simple analysis around use of Bayesian and deep learning packages as evidenced in the research literature and downloads to get a perspective on what actually happened and is happening. The comparison aims to approach the following goals with very simple research citation metrics:

1. How does Bayesian modeling software stack up against deep learning? 
2. Asses the impact of Bayesian modeling software using deep learning metrics as a yard stick--how big a fraction of a huge thing are we?
3. Contextualize the roles each approach has by looking at subject distributions. The technologies have very different use cases so one would expect variation.

Citation counting is a crude metric but it has the advantage of simplicity. In compiling these metrics I came away with very different opinion than I started with so I thought it worth sharing. My prior assumptions were that Bayesian modeling was very niche and scurrying around doing very useful and important science but niche none the less. This analysis led me to revise that opinion considerably.

For the purposes of this document we take Bayesian software to be Stan (@standev) and PyMC3 (@salvatier2016probabilistic) with ecosystem components included that are likely to be cited. That includes the simplified syntax interfaces to Stan, brms (@brms), RstanArm (@rstanarm), and interface packages to Stan, RStan (@rstan) and PyStan (@pystan)[^1]. The analysis applies only to packages that are in current development so the venerated, and very high impact, packages like BUGS (@lunn2013bugs) and JAGS (@plummer2003jags) are not considered although including them would double our counts. Also, the ecosystems are actually much larger but they are unlikely to be cited in the research literature and we have to stop some place. For deep learning we take TensorFlow, its interface Keras (@keras) and PyTorch (@pytorch) as roughly equivalent entities for the comparison. Theano (@theano) is no longer in development. It should be noted that both PyTorch and TensorFlow have implemented HMC with NUTS for Bayesian inference but those are recent developments that have not made much of an impact yet.

[^1]:There are also lighter weight interfaces as well interface to other languages that I ignore due to low expected research mentions, see [https://https://mc-stan.org/users/interfaces/]([https://https://mc-stan.org/users/interfaces/) for a listing.

The key resource behind this document is Elsevier's https://scopus.com research search engine that provides a tightly curated[^2] search index that includes sources outside of the academic behemoth's own journals. It also provides a solid API (Application Programmer Interface) and a classification of journals into subject areas. The actual form of Scopus queries is discussed below.

[^2]:https://scolar.google.com genrally yields double the counts for similar queries but no subject classification or API to code against. The increased counts are likely due to inclusion of non-peer reviewed sites like [https://arxiv.org](https://arxiv.org)

# High level prevalence of Deep Learning vs Bayesian Modeling

In the process of grant writing I create metrics to help justify projects, lately in partnership with PyMC and ArviZ through the scientific fiscal sponsor NumFOCUS of which Stan is a member as well. Since Bayesian modeling seems always in the shadow of deep learning I started tracking those software packages as well for comparison. Below we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with support/derivative packages RStan, RStanArm, PyStan and brms.

```{r Access Scopus, include=FALSE}
rm(list = rm()) #clean out any state from previous runs
library(tidyverse)
library(jsonlite)
library(httr)
library(kableExtra)
library(stringr)

# remember to set working directory to the location of this file
setwd("~/git/ScientificSoftwareImpactMetrics")
credentials <- read_json('Scopus_credentials.json')
API_KEY <- credentials$API_KEY
INSTITUTION_TOKEN <- credentials$INSTITUTION_TOKEN
# Format of Scopus_credentials.json
# {
# "API_KEY":"XXXXXXXXXXXXXXXXXXXXXXXXXxx",
# "INSTITUTION_TOKEN":"XXXXXXXXXXXXXXXXXXXXXXXX"
# }

BASE_URL = 'https://api.elsevier.com/content/search/scopus'

USE_CACHE = TRUE #will have to setup a redis server, strongly recommend doing this
REPORT_PROGRESS = FALSE

if (USE_CACHE) {
  library(redux)
  redis <- redux::hiredis()
  get_results <- function(url) {
    cache <- redis$GET(url) # check redis first
    if (!is.null(cache)) {
      result <- unserialize(cache)
      if (result$status_code == 200) {
        if (REPORT_PROGRESS) {
          cat(paste("\nhitting redis for", url))
        }
        return(result)
      }
    }
    random_wait <- abs(rnorm(1, 1, 1))
    if (REPORT_PROGRESS) {
      cat(paste("\ncache miss, querying:", url, "\n"))
      cat(paste(
        "\nWaiting",
        random_wait,
        "seconds to be nice to webserver\n"
      ))
    }
    Sys.sleep(random_wait)
    result <- GET(
      url,
      add_headers('X-ELS-APIKey' = API_KEY,
                  'X-ELS-Insttoken' = INSTITUTION_TOKEN)
    )
    redis$SET(url, serialize(result, NULL))
    return(result)
  }
} else { #not using cache, hacky but works with R
  get_results <- function(url) {
    random_wait <- abs(rnorm(1, 1, 1))
    if (REPORT_PROGRESS) {
      cat(paste("\nno cache used, querying:", url, "\n"))
      cat(paste(
        "\nWaiting",
        random_wait,
        "seconds to be nice to webserver\n"
      ))
    }
    Sys.sleep(random_wait)
    result <- GET(
      url,
      add_headers('X-ELS-APIKey' = API_KEY,
                  'X-ELS-Insttoken' = INSTITUTION_TOKEN)
    )
    return(result)
  }
}

# setup queries for analysis

stan_q <- '(gelman+AND+hoffman+AND+stan)+OR+mc-stan.org'
pymc_q <- 'PyMC3+OR+(PyMC*+AND+fonnesbeck)'
rstan_q <- '(rstan+AND+NOT+mit)'
brms_q <-  '(brms+AND+burkner)'

#Display names followed by queries
pkg_query <- c('Stan', 'brms','PyStan', 'RStanArm', 'RStan','PyMC',  'TensorFlow', 'PyTorch', 'Keras',
               stan_q, brms_q, 'pystan', 'rstanarm', rstan_q, pymc_q, 'tensorflow', 'pytorch', 'Keras')
pkg_query_m <- matrix(data = pkg_query, ncol =2)
queryBayVsDeepM <- matrix(ncol = 2, nrow = 2)
queryBayVsDeepM[1,] <- c('Bayesian', paste(pkg_query_m[1:6,2], collapse = "+OR+"))
queryBayVsDeepM[2,] <- c('Deep Learning', paste(pkg_query_m[7:9,2], collapse = "+OR+"))

```

```{r DL vs Bayes Graph, echo=FALSE, message=FALSE, warning=FALSE}
library(ggrepel)
year_start <- 2012
year_end <- 2020
years <- year_start:year_end
df <- data.frame(years)

years <- year_start:year_end
scopus.df <- data.frame(years)
for (i in 1:nrow(queryBayVsDeepM)) {
  package_name = queryBayVsDeepM[i, 1]
  query = queryBayVsDeepM[i, 2]
  if (query == '') {
    query = package_name
  }
  for (suffix in c('','_no_comp_sci')) {
    year_counts = c()
        for (year in year_start:year_end) {
      url <- paste(BASE_URL, '?query=','REF(', query,')',
                   "+AND+PUBYEAR+=+", year,
                   sep = '')
      if (suffix == '_no_comp_sci') {
        url = paste(url, '+AND+NOT+SUBJAREA(COMP)', sep = '')
      }
      result <- get_results(url)
      json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))
      data <- jsonlite::fromJSON(json_txt)
      year_counts = c(
        year_counts,
        as.numeric(data$`search-results`$`opensearch:totalResults`))
    }
    scopus.df[paste(
      package_name,
      suffix,
      sep = '')] <- year_counts
  }
}

column_names <- colnames(scopus.df)
col_sums <- colSums(scopus.df)
df_long <- gather(scopus.df, topic, yr_count,
                  column_names[2]:column_names[length(column_names)])

df_long_label <- df_long %>%
  mutate(label = if_else(years == max(years),
                         paste(as.character(topic),col_sums[topic], sep = ':'), NA_character_))

plot2 <-
  ggplot(data = df_long_label, aes(
    x = years,
    y = yr_count,
    group = topic,
    color = topic
  )) +
  geom_line() +
  geom_point() +
  geom_label_repel(aes(label = label),
                   na.rm = TRUE) +
  scale_color_discrete(guide = FALSE) +
  labs(y = "year count for research articles matching topic",
  caption = "Fig 1: Relative counts with and without computer science journals") +
   theme(plot.caption=element_text(size=12, hjust=0, margin=margin(15,0,0,0)))

print(plot2)
```

Reading from top of Fig 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,500--nearly a 10x difference. One could excuse a bit of Bayesian despondency in the face of the wave depicted above--this is ratio mirrors my expectations. But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 34,977 to 8,885--a 75% drop. That drop IS a surprise.  I didn't have a prior on deep learning's distribution across subject categories but I didn't expect it to be so skewed. The Bayesian ecosystem shows a total of 4,578 of which 14% are computer science--that drop does not surprise me.

An early take away is that Bayesian's shouldn't bother applying to computer science directorates since it's deep learning's home turf or Bayesians should double down on applying to computer science directorates since they must be getting sick of deep learning proposals.

# Subject Category Breakdowns

So what is going on here? Keeping speculation to a minimum, the next question is what does the the distribution look like for non-computer science subject areas?  Below is a table of subject counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed:

```{r Table 1, echo=FALSE, message=FALSE, warning=FALSE}
library(scales)
library(kableExtra)

makeDisplayTable <- function(df) {
  subjCats = colnames(df)[-1:-2]
  df %>% group_by(package) %>% 
    summarize_each(funs(sum), all_of(subjCats)) -> outDf
  finalDf = t(outDf)
  subFrame = finalDf[-1,]
  cleanDf =  as.data.frame(apply(subFrame, 2, as.numeric))
  colnames(cleanDf) = finalDf[1,]
  rownames(cleanDf) = rownames(finalDf)[-1]
  cleanDf %>% mutate(totals = rowSums(.)) -> summedDf
  displayDf = summedDf
  displayDf[] = NA_character_
  for (i in 1:nrow(summedDf)) {
    displayDf[i,] = as.list(sprintf("%d/%s", as.numeric(summedDf[i,]), 
                                    label_percent(accuracy = 1)(as.numeric(summedDf[i,]/summedDf[i,]$totals))))
    displayDf[i,]$totals = summedDf[i,]$totals 
  }
  displayOrderedDf <- displayDf[order(row.names(displayDf)),] 
  return(displayOrderedDf)
}

getAllSubjectCategories <- function (queryM) {
  year_start = 2006
  year_end = 2021
  years <- year_start:year_end
  accumDf = data.frame(matrix(ncol = 2, nrow = nrow(queryM) * length(years)))
  colnames(accumDf) = c('package', 'years')
  counter = 1
  article_count = 0
  article_subj_count = 0
  for (i in 1:nrow(queryM)) {
    package_name = queryM[i, 1]
    query = queryM[i, 2]
    if (!is.na(str_match(query,"^\\(")) && !is.na(str_match(query,"\\)$"))) {
      query = str_match(query, "^\\((.*)\\)$")[1,2] #remove extra brackets, breaks scopus
    }
    accumDf[counter:((counter - 1) + length(years)),]$years <- years
    accumDf[counter:((counter - 1) + length(years)),]$package <- 
      rep(package_name, length(years))
    for (year in year_start:year_end) {
      url <-
        paste(
          BASE_URL,
          '?query=',
          'REF(',
          query,
          ')',
          "+AND+PUBYEAR+=+",
          year,
          '&facets=subjarea(count=101)',
          sep = ''
        )
      result <- get_results(url)
      if (result$status_code != 200) {
        print(sprintf("got non 200 status from query: %d", result$status_code))
        stop()
      }
      json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))
      data <- jsonlite::fromJSON(json_txt)
      article_count <-
        as.numeric(data$`search-results`$`opensearch:totalResults`) + article_count
      facet_count <- length(data$`search-results`$facet$category$name)
      j <- 1
      while (j < facet_count) {
        name <- data$`search-results`$facet$category$label[j]
        name <- str_replace(name, " \\(all\\)", "")
        hitCount <- as.numeric(data$`search-results`$facet$category$hitCount[j])
        article_subj_count <- hitCount + article_subj_count
        if (!name %in% colnames(accumDf)) {
          accumDf[name] <- rep(0, length(years) * nrow(queryM))
          #print(paste("name=",name,", count=",hitCount))
        }
        accumDf[counter,][[name]] <- hitCount
        j <- j+ 1
      }
      counter = counter + 1
    }
  }
  if (REPORT_PROGRESS) {
    print(sprintf("%d paperXsubject with %d total papers for %.3f avg subjects per paper",
                  article_subj_count,
                  article_count,
                  article_subj_count/article_count))
  }
  return(accumDf)
}

bayesVdeepLearnDf <- getAllSubjectCategories(queryBayVsDeepM)
displayBayesVdeepLearnDf <- makeDisplayTable(bayesVdeepLearnDf)

detailDf  <- getAllSubjectCategories(pkg_query_m)
displayDetailDf <- makeDisplayTable(detailDf)

allDataDf <- cbind(displayBayesVdeepLearnDf,displayDetailDf)

kable(allDataDf, booktabs = TRUE) %>% kable_styling(font_size = 10)

```

Table 1 shows the basic subject split between Bayesian modeling and deep learning with total for the two categories, following are the individual package counts with a total as well. Note that the totals do not match because an article can match more than one package and as result gets contributes to the count of multiple packages. On average an article's journal is classified into two subject categories.

Staring with the subject that got us here we see that computer science has 728 Bayesian publications vs 30,029 deep learning publications for 98% of the split. The left most two columns are the same queries used for the graph in Fig 1 with the total in column 3. Columns 4-8 are indivudual packages with percentage of the total in the 9th column.  One sees that TensorFlow has 18,178 references with 51% of the count--note that the total counts are higher because an article can matched by more that one query, each package gets a query, so the article count is 3 if Keras, PyMC and TensorFlow queries match the references and it has a computer science subject classification.

Looking at the tea leaves I leave to others but I'll make the observation that Bayesians are playing a pretty big game in a lot of areas of science and as someone interested in getting funding to Bayesian software this is a pretty strong argument that we are highly relevant.

# Details of the queries

Below are the queries used to identify the packages. I would estimate a 5% false positive rate based based on informal examination of return sets. For example the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released. The name 'Stan' presented obvious difficulties and the combination shown achieved desired count drop off at 2012 with reasonable precision and each of the 'OR' terms contributed to coverage significantly. How the word 'mc-stan.org' was tokenized is unknown but there were no results before 2013 for the bare query.

```{r Query Table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
pkg_query_m_display <- pkg_query_m
pkg_query_m_display[,2] <- sprintf("REF(%s)", str_replace_all(pkg_query_m[,2], '\\+',' '))

kable(pkg_query_m_display, booktabs = TRUE) %>% kable_styling(font_size = 12)

```

PyMC presented some challenges in that the string PyMC was referenced in early publications but also is ambiguous with chemical terms, so like Stan, author names were included to restrict search as was the case with 'brms'. These queries were also developed searching entire documents rather than the current restriction to the references section of papers. I would expect the false positive rate to be even lower.

## Technical details and access to source/scripts

Supporting technology for this document includes the R language, the tidyverse group of packages that give data and visualization tools and finally this Rmarkdown document developed using the Rstudio IDE (Integrated Development Environment). I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/bayesianImpactMetrics](https://github.com/breckbaldwin/bayesianImpactMetrics) where it lives as `index.Rmd`, is rendered as `index.html` and is viewable as [https://breckbaldwin.github.io/bayesianImpactMetrics/](https://breckbaldwin.github.io/bayesianImpactMetrics/). All source code is viewable in the .Rmd document.

### 2021-05-23 DeepLearningAndBayesianSoftware.Rmd
getting close to release

@@
 ---
-title: "Deep Learning and Bayesian Modeling: Comparing research use via citation counting"
+title: "Deep Learning and Bayesian Modeling"
+subtitle: "Comparing research use via citation counting"
 author: "Breck Baldwin"
-date: "5/19/2021"
+email: "breckbaldwin@gmail.com"
+date: "5/23/2021"
 output:
- bookdown::tufte_html2: default
+ bookdown::html_document2:
+   includes:
+     in_header: _html/ga.html 
 bibliography: citations.bib
@@
 
+# Introduction
+
 Stan, a Bayesian modeling language, was released in 2012 to considerable fanfare.
@@
 Yet Bayesians found themselves in deep learning's shadow somehow.
-It is 2021 and this document does some very simple analysis around use of Bayesian and deep learning packages as evidenced in the research literature and downloads to get a perspective on what actually happened and is happening.
+It is 2021 and this document does some very simple analysis around use of Bayesian and deep learning packages as evidenced in the research literature to get a perspective on what actually happened and is happening.
 The comparison aims to approach the following goals with very simple research citation metrics:
@@
 So what is going on here?
-Keeping speculation to a minimum, the next question is what does the the distribution look like for non-computer science subject areas?
+The next question is what does the the distribution look like for non-computer science subject areas?
 Below is a table of subject counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed:
@@
 allDataDf <- cbind(displayBayesVdeepLearnDf,displayDetailDf)
-
+allDataDf <- allDataDf[,c(1,2,3,11,6,7,10,9,4,8,5,12,13)]
+displayedColNames <- colnames(allDataDf)
+colnames(allDataDf) <- c(displayedColNames[1:12],'detail totals')
 kable(allDataDf, booktabs = TRUE) %>% kable_styling(font_size = 10)
@@
 
-Table 1 shows the basic subject split between Bayesian modeling and deep learning with total for the two categories, following are the individual package counts with a total as well.
+Table 1 shows the basic subject split between Bayesian modeling and deep learning with total for the two categories, following are the individual package counts with a 'detail totals' as well for the individual packages.
 Note that the totals do not match because an article can match more than one package and as result gets contributes to the count of multiple packages.
@@
 
-Staring with the subject that got us here we see that computer science has 728 Bayesian publications vs 30,029 deep learning publications for 98% of the split.
+Staring with the subject that got us here we see that computer science has 728 Bayesian publications vs 29,301 deep learning publications for 98% of the share of 30,029 total.
 The left most two columns are the same queries used for the graph in Fig 1 with the total in column 3.
-Columns 4-8 are indivudual packages with percentage of the total in the 9th column.
-One sees that TensorFlow has 18,178 references with 51% of the count--note that the total counts are higher because an article can matched by more that one query, each package gets a query, so the article count is 3 if Keras, PyMC and TensorFlow queries match the references and it has a computer science subject classification.
+Columns 4-12 are individual packages with percentage of the total in the 13th column.
+Sticking with computer science, one sees that TensorFlow has 18,178 references with 51% of the count--note that the total counts are higher because an article can matched by more that one query, each package gets a query, so the article count is 3 if Keras, PyMC and TensorFlow queries match the references and it has a computer science subject classification.
 
-Looking at the tea leaves I leave to others but I'll make the observation that Bayesians are playing a pretty big game in a lot of areas of science and as someone interested in getting funding to Bayesian software this is a pretty strong argument that we are highly relevant.
+I leave the tea leaf interpretation to others but I'll make the observation that Bayesians are playing a pretty big game in a lot of areas of science and as someone interested in getting funding to Bayesian software this is a pretty strong argument that we are highly relevant.
 
@@
 Below are the queries used to identify the packages.
+Direct citation linking does not work well with these packages since many of them had no journal publication to cite in a referring article.
+As a result a string search was used to identify the citations in the reference section as determined by Scopus.com's indexing engine.
 I would estimate a 5% false positive rate based based on informal examination of return sets.
-For example the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released.
+
+For an example of how the queries were developed, the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released.
 The name 'Stan' presented obvious difficulties and the combination shown achieved desired count drop off at 2012 with reasonable precision and each of the 'OR' terms contributed to coverage significantly.
-How the word 'mc-stan.org' was tokenized is unknown but there were no results before 2013 for the bare query.
+How the word 'mc-stan.org' was tokenized is unknown but there were no results before 2013 for the solo query.
 
@@
 PyMC presented some challenges in that the string PyMC was referenced in early publications but also is ambiguous with chemical terms, so like Stan, author names were included to restrict search as was the case with 'brms'.
-These queries were also developed searching entire documents rather than the current restriction to the references section of papers.
+These queries were initially developed to search entire documents rather than the current restriction to the references section of papers.
 I would expect the false positive rate to be even lower.
 
-## Technical details and access to source/scripts
+# Access to source/scripts and acknowledgements
 
-Supporting technology for this document includes the R language, the tidyverse group of packages that give data and visualization tools and finally this Rmarkdown document developed using the Rstudio IDE (Integrated Development Environment).
-I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/bayesianImpactMetrics](https://github.com/breckbaldwin/bayesianImpactMetrics) where it lives as `index.Rmd`, is rendered as `index.html` and is viewable as [https://breckbaldwin.github.io/bayesianImpactMetrics/](https://breckbaldwin.github.io/bayesianImpactMetrics/).
+Supporting technology for this document includes the R language (@Rcite), the 'tidyverse' group of packages (@tidyverse) that give data and visualization tools and this Rmarkdown document (@rmarkdown) rendered with 'bookdown' (@bookdown) developed using the Rstudio (@rstudio) IDE (Integrated Development Environment).
+Supporting package include 'kable-extra' (@kable-extra), a table formatter, the 'scales' (@scales) package for converting to percentages, 'ggrepel' (@ggrepel) for labeling line graphs with dynamically shifting labels, 'jsonlite' (@jsonlite) for JSON data parsing, 'httr' (@httr) for GET requests, 'stringr' (@stringr) for regular expression matching and 'redux' (@redux) for access to the Redis server (@redis) which I highly recomend using as a caching layer for this sort of project.
+
+I'd like to thank the Andrew Gelman's Laplace lab at Columbia University for support.
+
+I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/bayesianImpactMetrics](https://github.com/breckbaldwin/bayesianImpactMetrics) where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as [https://breckbaldwin.github.io/bayesianImpactMetrics/DeepLearningAndBayesianSoftware.html](https://breckbaldwin.github.io/bayesianImpactMetrics/DeepLearningAndBayesianSoftware.html).
 All source code is viewable in the .Rmd document.
+
+I can be reached at breckbaldwin@gmail.com or via the github issues on the repository.
+
+# References

### 2021-05-23 DeepLearningAndBayesianSoftware.Rmd
hopefully final

@@
 
-I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/bayesianImpactMetrics](https://github.com/breckbaldwin/bayesianImpactMetrics) where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as [https://breckbaldwin.github.io/bayesianImpactMetrics/DeepLearningAndBayesianSoftware.html](https://breckbaldwin.github.io/bayesianImpactMetrics/DeepLearningAndBayesianSoftware.html).
+I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics](https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics) where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as [https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html](https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html).
+Note that github pages changes urls around in odd ways for the rendered view of hte html page.
 All source code is viewable in the .Rmd document.

### 2021-05-23 DeepLearningAndBayesianSoftware.Rmd
minor edits, added a stub for conclusions

@@
 And then deep learning happened.
-The relese of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to everyone and in 2016 began garnering significant traction.
+The release of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to everyone and in 2016 began getting significant traction.
 One result of this was people often assumed that deep learning's successes usurped Bayesian modeling's domain.
@@
 Yet Bayesians found themselves in deep learning's shadow somehow.
+
 It is 2021 and this document does some very simple analysis around use of Bayesian and deep learning packages as evidenced in the research literature to get a perspective on what actually happened and is happening.
@@
 In the process of grant writing I create metrics to help justify projects, lately in partnership with PyMC and ArviZ through the scientific fiscal sponsor NumFOCUS of which Stan is a member as well.
-Since Bayesian modeling seems always in the shadow of deep learning I started tracking those software packages as well for comparison.
+Since Bayesian modeling seems always in the shadow of deep learning I started tracking deep learning software packages as well for comparison.
 Below we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with support/derivative packages RStan, RStanArm, PyStan and brms.
@@
 Reading from top of Fig 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,500--nearly a 10x difference.
-One could excuse a bit of Bayesian despondency in the face of the wave depicted above--this is ratio mirrors my expectations.
+One could excuse a bit of Bayesian despondency in the face of the wave depicted above--this ratio supports my expectations.
 But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 34,977 to 8,885--a 75% drop.
@@
 The Bayesian ecosystem shows a total of 4,578 of which 14% are computer science--that drop does not surprise me.
-
-An early take away is that Bayesian's shouldn't bother applying to computer science directorates since it's deep learning's home turf or Bayesians should double down on applying to computer science directorates since they must be getting sick of deep learning proposals.
 
@@
 Direct citation linking does not work well with these packages since many of them had no journal publication to cite in a referring article.
-As a result a string search was used to identify the citations in the reference section as determined by Scopus.com's indexing engine.
+As a result a string search was used to identify the citations in the reference section which was determined by Scopus.com's document parser.
 I would estimate a 5% false positive rate based based on informal examination of return sets.
@@
 For an example of how the queries were developed, the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released.
-The name 'Stan' presented obvious difficulties and the combination shown achieved desired count drop off at 2012 with reasonable precision and each of the 'OR' terms contributed to coverage significantly.
+The name 'Stan' presented obvious difficulties, see [https://statmodeling.stat.columbia.edu/2019/04/29/we-shouldntve-called-it-stan-i-shouldve-listened-to-bob-and-hadley/](https://statmodeling.stat.columbia.edu/2019/04/29/we-shouldntve-called-it-stan-i-shouldve-listened-to-bob-and-hadley/).
+The query shown achieved desired count drop off at 2012 with reasonable precision and each of the 'OR' terms contributed to coverage significantly.
 How the word 'mc-stan.org' was tokenized is unknown but there were no results before 2013 for the solo query.
@@
 I would expect the false positive rate to be even lower.
+
+# Conclusions
+
+I am waiting on some feedback from the Stan community before making any observations. But I will say that this overall has me feeling quite optimistic about the future of Bayesian modeling.
 

### 2021-05-24 DeepLearningAndBayesianSoftware.Rmd
added in emcee hammer, rejected inclusion in the end, too specialized

@@
 The analysis applies only to packages that are in current development so the venerated, and very high impact, packages like BUGS (@lunn2013bugs) and JAGS (@plummer2003jags) are not considered although including them would double our counts.
+There is also the best named Bayesian package of all time, "emcee: The MCMC Hammer", (@foreman2013emcee) which enjoys tremendous use in the astrophysics community but is too specialized to be considered a general Bayesian package so it is not included.
 Also, the ecosystems are actually much larger but they are unlikely to be cited in the research literature and we have to stop some place.
@@
     )
+    if (result$status_code != 200) { 
+      stop(sprintf("Got error from search, got status code %d",result$status_code))
+    }
     redis$SET(url, serialize(result, NULL))
@@
     )
+    if (result$status_code != 200) { 
+      stop(sprintf("Got error from search, got status code %d",result$status_code))
+    }
     return(result)
@@
 pymc_q <- 'PyMC3+OR+(PyMC*+AND+fonnesbeck)'
-rstan_q <- '(rstan+AND+NOT+mit)'
-brms_q <-  '(brms+AND+burkner)'
+rstan_q <- 'rstan+AND+NOT+mit'
+brms_q <-  'brms+AND+burkner'
+emcee_q <- 'emcee+AND+foreman-mackey'
 
 #Display names followed by queries
-pkg_query <- c('Stan', 'brms','PyStan', 'RStanArm', 'RStan','PyMC',  'TensorFlow', 'PyTorch', 'Keras',
-               stan_q, brms_q, 'pystan', 'rstanarm', rstan_q, pymc_q, 'tensorflow', 'pytorch', 'Keras')
-pkg_query_m <- matrix(data = pkg_query, ncol =2)
+pkg_query <- c('emcee', 'Stan', 'brms','PyStan', 'RStanArm', 'RStan','PyMC',  'TensorFlow', 'PyTorch', 'Keras',
+               emcee_q, stan_q, brms_q, 'pystan', 'rstanarm', rstan_q, pymc_q, 'tensorflow', 'pytorch', 'Keras')
+pkg_query_m <- matrix(data = pkg_query[c(-1,-11)], ncol =2) #emcee removed
 queryBayVsDeepM <- matrix(ncol = 2, nrow = 2)
-queryBayVsDeepM[1,] <- c('Bayesian', paste(pkg_query_m[1:6,2], collapse = "+OR+"))
-queryBayVsDeepM[2,] <- c('Deep Learning', paste(pkg_query_m[7:9,2], collapse = "+OR+"))
+queryBayVsDeepM[1,] <- c('Bayesian', 
+                         paste0('(', paste(pkg_query_m[1:6,2], collapse = ")+OR+("),')'))
+queryBayVsDeepM[2,] <- c('Deep Learning', 
+                         paste0('(', paste(pkg_query_m[7:9,2], collapse = ")+OR+("), ')'))
 
@@
 scopus.df <- data.frame(years)
-for (i in 1:nrow(queryBayVsDeepM)) {
-  package_name = queryBayVsDeepM[i, 1]
-  query = queryBayVsDeepM[i, 2]
-  if (query == '') {
-    query = package_name
-  }
+#queryM = pkg_query_m
+queryM = queryBayVsDeepM
+for (i in 1:nrow(queryM)) {
+  package_name = queryM[i, 1]
+  query = queryM[i, 2]
   for (suffix in c('','_no_comp_sci')) {
+#  for (suffix in c('')) {
     year_counts = c()
@@
 
+   log_plot = plot2 + 
+     scale_y_continuous(breaks=c(1, 10, 100, 1000, 10000, 100000, 1e+06, 1e+07, 1e+08), 
+                     trans = scales::log_trans())
 print(plot2)
@@
     query = queryM[i, 2]
-    if (!is.na(str_match(query,"^\\(")) && !is.na(str_match(query,"\\)$"))) {
-      query = str_match(query, "^\\((.*)\\)$")[1,2] #remove extra brackets, breaks scopus
-    }
     accumDf[counter:((counter - 1) + length(years)),]$years <- years
@@
       result <- get_results(url)
-      if (result$status_code != 200) {
-        print(sprintf("got non 200 status from query: %d", result$status_code))
-        stop()
-      }
       json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))

### 2021-05-24 DeepLearningAndBayesianSoftware.Rmd
added in Oriol thanks, added the group queries, typo fixes

@@
 
-Reading from top of Fig 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,500--nearly a 10x difference.
+Reading from top of Fig 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,600--nearly a 10x difference.
 One could excuse a bit of Bayesian despondency in the face of the wave depicted above--this ratio supports my expectations.
-But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 34,977 to 8,885--a 75% drop.
+But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 35k to 8.8k--a 75% drop.
 That drop IS a surprise.
 I didn't have a prior on deep learning's distribution across subject categories but I didn't expect it to be so skewed.
-The Bayesian ecosystem shows a total of 4,578 of which 14% are computer science--that drop does not surprise me.
+The Bayesian ecosystem shows a total of 4.5k of which 14% are computer science--that ratio does not surprise me.
 
@@
 
-Staring with the subject that got us here we see that computer science has 728 Bayesian publications vs 29,301 deep learning publications for 98% of the share of 30,029 total.
+Staring with the subject that got us here we see that computer science has 727 Bayesian publications vs 29,365 deep learning publications for 98% of the share of 30,092 total.
 The left most two columns are the same queries used for the graph in Fig 1 with the total in column 3.
@@
 ```{r Query Table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-pkg_query_m_display <- pkg_query_m
-pkg_query_m_display[,2] <- sprintf("REF(%s)", str_replace_all(pkg_query_m[,2], '\\+',' '))
+pkg_query_m_display <- rbind(pkg_query_m, queryBayVsDeepM)
+pkg_query_m_display[,2] <- sprintf("REF(%s)", str_replace_all(pkg_query_m_display[,2], '\\+',' '))
 
@@
 I would expect the false positive rate to be even lower.
+
+The above queries can be run with the advance search option at https://scopus.com and are generated from the API queries actually used.
+The change is replacing '+' with ' '.
+An account is required which is included for many education institutions.
 
@@
 
-I'd like to thank the Andrew Gelman's Laplace lab at Columbia University for support.
+I'd like to thank Andrew Gelman and the Laplace lab at Columbia University for support.
+
+Oriol Abril Pla brought up emcee as a popular Bayesian system worth consideration which was a big oversight on my part as well as providing insightful commentary.
 

### 2021-08-26 DeepLearningAndBayesianSoftware.Rmd
submitted to journal

@@
 ---
-title: "Deep Learning and Bayesian Modeling"
+title: "Deep Learning does not Replace Bayesian Modeling"
 subtitle: "Comparing research use via citation counting"
@@
 email: "breckbaldwin@gmail.com"
-date: "5/23/2021"
+date: "8/20/2021"
 output:
@@
      in_header: _html/ga.html
+
 bibliography: citations.bib
+csl: american-medical-association.csl
 link-citations: yes
@@
 And then deep learning happened.
-The release of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to everyone and in 2016 began getting significant traction.
+The release of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to all and by it 2016 began getting significant traction.
 One result of this was people often assumed that deep learning's successes usurped Bayesian modeling's domain.
-This was not in our imagination, NSF reviews came back dismissing Stan funding because all the interesting work was assumed to be happening with deep learning.
-Mind you, deep learning and Bayesian modeling are conceptual cousins but very different from each other and are better thought of as complementary than as antagonistic.
+This is not in our collective imagination--Bayesians like to 'believe things' after all--NSF reviews came back dismissing Stan funding because all the interesting work was assumed to be happening with deep learning.
+Recently in the UK, open skepticism was expressed about impact claims for Bayesian software in response to a research grant.
+Mind you, deep learning and Bayesian modeling are conceptual cousins but in the end are very different from each other.
+They are better thought of as complementary than as antagonistic.
 Yet Bayesians found themselves in deep learning's shadow somehow.
@@
 
-1. How does Bayesian modeling software stack up against deep learning? 
+1.
+How does Bayesian modeling software stack up against deep learning without appeal to feature comparisons, performance arguments on suspect data sets or achieved closeness to Platonic ideals?
+Just go out and count how many citations the respective approaches have in the research literature.
+Counting and categorization, that's it.
 2. Asses the impact of Bayesian modeling software using deep learning metrics as a yard stick--how big a fraction of a huge thing are we?
@@
 
-[^1]:There are also lighter weight interfaces as well interface to other languages that I ignore due to low expected research mentions, see [https://https://mc-stan.org/users/interfaces/]([https://https://mc-stan.org/users/interfaces/) for a listing.
+[^1]:There are also lighter weight interfaces as well interface to other languages that I ignore due to low expected research mentions, see https://https://mc-stan.org/users/interfaces/ for a listing.
 
@@
 It also provides a solid API (Application Programmer Interface) and a classification of journals into subject areas.
-The actual form of Scopus queries is discussed below.
+The actual form of Scopus queries is discussed below which should allow the questioning reader to verify and update the counts for the various packages discussed.
 
 [^2]:https://scolar.google.com genrally yields double the counts for similar queries but no subject classification or API to code against.
-The increased counts are likely due to inclusion of non-peer reviewed sites like [https://arxiv.org](https://arxiv.org)
+The increased counts are likely due to inclusion of non-peer reviewed sites like https://arxiv.org
+
+<br>
 
@@
 For an example of how the queries were developed, the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released.
-The name 'Stan' presented obvious difficulties, see [https://statmodeling.stat.columbia.edu/2019/04/29/we-shouldntve-called-it-stan-i-shouldve-listened-to-bob-and-hadley/](https://statmodeling.stat.columbia.edu/2019/04/29/we-shouldntve-called-it-stan-i-shouldve-listened-to-bob-and-hadley/).
+The name 'Stan' presented obvious difficulties, see https://statmodeling.stat.columbia.edu/2019/04/29/we-shouldntve-called-it-stan-i-shouldve-listened-to-bob-and-hadley/.
 The query shown achieved desired count drop off at 2012 with reasonable precision and each of the 'OR' terms contributed to coverage significantly.
@@
 
-The above queries can be run with the advance search option at https://scopus.com and are generated from the API queries actually used.
-The change is replacing '+' with ' '.
+The above queries can be run with the advance search option at https://scopus.com and are generated from the API queries actually used where '+' was replaced with ' ' above.
 An account is required which is included for many education institutions.
@@
 
-I am waiting on some feedback from the Stan community before making any observations. But I will say that this overall has me feeling quite optimistic about the future of Bayesian modeling.
+As often happens conducting a closer look reveals subtleties.
+Deep learning dominates the computer science part of science and obviously it dominates in the commercial sector which this article does not address.
+However the remaining parts of science show a more balanced distribution across the approaches.
+This makes sense since the approaches are very different and serve different goals.
+
+As a Bayesian advocate I'd like to leave the funding agencies with the observation that Bayesian software carries roughly one third of the research load by research citation count (3925/(3925+8884)) without the backing of Fortune 500 companies like Facebook and Google.
+Just recently I learned of a funding denial because the funders didn't believe that Bayesian software could have the impact that was being claimed--that opinion could well be self fulfilling in the future but as of now it is not justified.
 
@@
 
-I intend this to be an evolving document with periodic updates at it's github repo at [https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics](https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics) where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as [https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html](https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html).
+I intend this to be an evolving document with periodic updates at it's github repo at https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html.
 Note that github pages changes urls around in odd ways for the rendered view of hte html page.

### 2021-10-16 Author Response Letter.Rmd
backup commit, drafts not ready

---
title: "Response to Reviewers for: Deep Learning does not Replace Bayesian Modeling"
author: "Breck Baldwin"
email: "breckbaldwin@gmail.com"
date: "10/15/2021"
output:
 bookdown::html_document2:
   includes:
     in_header: _html/ga.html

bibliography: citations.bib
csl: american-medical-association.csl
link-citations: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	comment = NA
)
```

``` {echo=TRUE}
Reviewer: 1

Comments to the Author
An interesting paper with much to think on. It is noticeable that Deep Learning is now the chosen method in most fields even though it is computer science which dominates. This should perhaps be noted. It would be interesting to consider whether adopting Deep Learning in these fields is always appropriate or if Bayesian methods could offer a more appropriate alternative.
```

Author Response

R1:
This paper also demonstrates that deep learning is not the default tool when it is time for some reasoning in the scientific enterprise. If you are doing immunology/microbiology then Bayesian methods dominate. Environmental science, economics are a coin flip. While the choice of machine reasoning techniques should be entirely determined by the problem being addressed, researchers use what is popular and being talked about. Perhaps this paper gets Bayesian methods considered a bit more. 
:R1

Added comment

``` {echo=TRUE}
Reviewer: 2

Comments to the Author
The author provides a robust and necessary analysis of the interest around and funding dedicated to the areas of Bayesian inference and Deep Learning. The analysis yields interesting insights into the utilisation of these learning frameworks in a variety of fields. The article is concise, well written and I in my opinion is ready for publication as is.
```

``` {echo=TRUE}
Reviewer: 3

Comments to the Author
A thought-provoking paper, which with some refinement provides further evidence regarding the importance of computational statistics - often peddled as AI hype.

I have a series of questions & comments that I would welcome author response to.

Abstract: you comment “once computer science is subtracted”. How can you argue for this though, given that so much work in both deep learning and probabilistic modeling (though as below, please don’t assume these are different!) sits squarely in computer science! This feels, without full argument, like cherry-picking - and of course, in the subjects that are left (dominated by Stats & Applied Math) it’s inevitable that we find more usage of Bayesian methods (and probably more use of R as opposed to Python too!)

Google, Facebook, Microsoft and the gang are strong exponents of Bayes! Let’s remember that Bayesian Optimization sits at the core of hyper-parameter tuning for AlphaGoZero and (behind the scenes) for much model updating in Facebook.
```
### Abstract revised:

One could be excused for assuming that deep learning had or will soon usurp all credible work in reasoning, artificial intelligence and statistics, but like most 'meme' class broad generalizations the concept does not hold up to scrutiny. Memes don't generally matter since the experts will always know better but in the case of Bayesian software like Stan and PyMC3 even its developers and advocates bemoan the apparent dominance of deep learning as manifested in popular culture, breathtaking performance and most problematically from funding agency peer review that impacts our ability to further advance the field. The facts however do not support the assumed dominance of deep learning in science upon closer examination.

This letter simply makes the argument by the crudest of possible metrics, citation count, that once R2:the discipline of:R2 Computer Science is subtracted, Bayesian software accounts for nearly a third of research citations. Stan and PyMC3 dominate some fields, PyTorch, Keras and TensorFlow dominate others with lots of variation in between. Bayesian and deep learning approaches are related but very different technologies in goals, implementation and applicability with little actual overlap so this is not a surprise, R3: e.g., deep learning cannot bring the explainability of applied math/statistics and Bayesian methods do not scale to deep learning data sets.

While deep learning behemoths like Facebook and Google both use and support Bayesian efforts, the Bayesian packages scientists actually use are academic/volunteer efforts punching far above their weight class and they need financial support. It would behoove funders to fully understand the impact and role of Bayesian methods in resource allocation.:R3

```{echo=TRUE}

Introduction: There is a *huge* rise in the use of Bayesian Deep learning. Please do not fall into the trap of making DL look like some ‘other’ thing. It’s just a non/hyper-parametric model and as such we can do all the full Bayesian stuff - and many do! There are well-documented works on semi-structured HMC, Riemann MC and good old fashioned MCMC for use in big deep nets, let alone all the neat work on approximate Bayes, mainly using variational learning. Look at work on things like loss-calibrated Bayesian deep nets, Bayes DL, Bayesian autoencoding, ‘HamilTorch’ MCMC package for TensorFlow (https://adamcobb.github.io/journal/hamiltorch.html) etc. You just can’t ignore all this. (and please don’t forget all the fine folks doing {S,O,P}DiffEq learning using things like graph NNs and ResNets - lots of Bayes to please everyone in the ‘physics meets DL’ world).
```

Intro: you claim NSF dismissal of (eg) Stan - where is the evidence citation for this?

Intro: Bayesian software - myopic - more to Bayes than sampling (which is non-Bayesian in any case!). Lots of work in AI using approximate Bayes, ABC, VB and so on. Gaussian Processes are HUGE in their own right too - but no MCMC in sight normally. You really need to note that Gaussian Processes, and Neural Kernel methods, are *fully Bayesian* yet DL replacements for many applications.

Sect 3. I just do not buy that packages are a proxy for research work. As before, what about all the "Bayesian" dropout work in NNs etc. VAEs? BayesDL tends not to use packages.

So, this has the making of a really interesting commentary - but really needs to address these issues before it can be published. I feel you do Bayes in DL a disservice by commenting just on big package usage.
```

Reviewer: 4

Comments to the Author
The author presents a lightweight but thought-provoking exposition of the scientific literature, providing interesting evidence that can be used by grantees and funders alike to justify funding statistical science (in particular Bayesian statistics) alongside broader computer science techniques badged under "AI". I was slightly surprised not to see references to Bayesian deep learning, or work at the intersection of Bayesian statistics and deep learning (e.g. using DNN as an approximation to a likelihood function), to support the point around complementarity, but this is a minor point and not worth a revision to the paper. I do hope the paper provokes further debate on this important topic.

#

### 2021-10-16 DeepLearningAndBayesianSoftware_changes_noted.Rmd
backup commit, drafts not ready

--- 2021-08-26 DeepLearningAndBayesianSoftware.Rmd
@@
 How does Bayesian modeling software stack up against deep learning without appeal to feature comparisons, performance arguments on suspect data sets or achieved closeness to Platonic ideals?
-Just go out and count how many citations the respective approaches have in the research literature.
+My response is--just go out and count how many citations the respective approaches have in the research literature.
 Counting and categorization, that's it.
@@
 This analysis led me to revise that opinion considerably.
+
+R4: 
+## Characterizing the difference between Stan/PyMC3 and TensorFlow/PyTorch
+
+What, if anything, are the hard differences between Bayesian modeling and deep learning?
+Deep learning can be implemented in Bayesian modeling, and lots of Bayesian concepts get used in the deep learning world as one reviewer noted:
+
+"There is a *huge* rise in the use of Bayesian Deep learning.
+Please do not fall into the trap of making DL look like some ‘other’ thing.
+It’s just a non/hyper-parametric model and as such we can do all the full Bayesian stuff - and many do!
+There are well-documented works on semi-structured HMC, Riemann MC and good old fashioned MCMC for use in big deep nets, let alone all the neat work on approximate Bayes, mainly using variational learning.
+Look at work on things like loss-calibrated Bayesian deep nets, Bayes DL, Bayesian autoencoding, ‘HamilTorch’ MCMC package for TensorFlow (https://adamcobb.github.io/journal/hamiltorch.html) etc."
+
+Perhaps Stan and PyMC3 are really part of the same discipline as deep learning and this article catalogs usage differences of siblings from the same parents--sort of like the difference between R and Python?
+
+But no.
+Another reviewer comment contains the key insight: they said it was no surprise that Bayesian software had high usage outside of computer science since other subjects are "dominated by Stats & Applied Math".
+I took this to mean that deep learning is not a natural candidate for use in those fields, and software like Stan and PyMC3 are.
+Unpacking that a bit I'll observe:
+
+* Unless the phenomenon under study is a physical neural net, deep learning only offers prediction services.
+While outstanding progress has been made and further progress may well involve Bayesian concepts, it is only prediction in the end.
+* The basis of prediction in deep learning is opaque to human comprehension even if Bayesian techniques are used. This applies to generative neural nets as well. 
+* Opacity blocks use in fields "dominated by Stats & Applied Math" where the goal involves developing and fitting mechanistic models.
+The science is in the model description, quality of fit validates the model.
+A high quality fit in the absence of an understandable model does not help.
+
+Deep learning clearly is used in mechanistic models but typically as a sensor or classifier, e.g., classify x-ray images for evidence of COVID pneumonia in an epidemialogical study.
+That study itself will likely be a state-based model where transition rates are explicitly estimated statistically and are human interpretable.
+
+Another case are hybrid models where a deep learning component replaces some or all of the likelihood of a Bayesian model.
+While the overall model may be interpretable, the individual deep learning components remain black boxes.
+Interpretability comes from have well understood properties from how they were trained etc.
+For example Maggie Lieu's presentation at StanCon that used a deep learning component to substitute for a numerically unstable galaxy mass estimator().
+
+Also Bayesian models can be utterly opaque if not authored with an eye to understandability, but the option for understandability exists and is generally the expectation.
+Deep learning systems cannot match this because a big portion of 'authoring' (hyper parameter setting) in those frameworks are beyond human understanding.
+
+In the game of "one of these things is not like the other", human interpretable models distinguishes Stan and PyMC3 from PyTorch and TensorFlow in quite absolute terms[^3].
+
+If I am correct that deep learning and Bayesian modeling are so different when it comes to interpretablity then the mingling of the fields baffles the mind.
+They share similar programming techniques, have similar research leading edges and ...
+.
+I think Stan got a hold of the AI tail which whipped it out of booring stats into ML.
+
+[^3]:However, in a year this article may be much more difficult to write since both projects have adopted HMC implementations but I'd suggest that is more of a grafting of another approach to their code base rather than natural growth of deep learning techniques into the PyMC3/Stan space.
+
+:R4
+
+## Defining what counts as a citation
 
@@
 
+R1: This paper also demonstrates that deep learning is not neccessarily the default tool when it is time for some reasoning in the scientific enterprise.
+If you are doing immunology/microbiology then Bayesian methods dominate.
+Environmental science, economics are a coin flip.
+Researchers use what is fashionable perhaps as a default--it may ease publication, be more familiar to students/postdocs but at the potential cost of being a bad fit for the problem at hand.
+If Bayesian methods are good enough for immunologists then perhaps other fields should have a look.:R1
+
 As a Bayesian advocate I'd like to leave the funding agencies with the observation that Bayesian software carries roughly one third of the research load by research citation count (3925/(3925+8884)) without the backing of Fortune 500 companies like Facebook and Google.

### 2021-11-05 DeepLearningAndBayesianSoftware.Rmd
final

@@
 email: "breckbaldwin@gmail.com"
-date: "8/20/2021"
+date: "11/01/2021"
 output:
@@
 
+<!--**Abstract revised: Note this is not in the paper but a separate attachment**
+
+One could be excused for assuming that deep learning had or will soon usurp all credible work in reasoning, artificial intelligence and statistics, but like most 'meme' class broad generalizations the concept does not hold up to scrutiny.
+Memes don't generally matter since the experts will always know better but in the case of Bayesian software like Stan and PyMC3 even its developers and advocates bemoan the apparent dominance of deep learning as manifested in popular culture, breathtaking performance and most problematically from funding agency peer review that impacts our ability to further advance the field.
+The facts however do not support the assumed dominance of deep learning in science upon closer examination.
+
+This letter simply makes the argument by the crudest of possible metrics, citation count, that once the discipline of Computer Science is subtracted, Bayesian software accounts for nearly a third of research citations.
+Stan and PyMC3 dominate some fields, PyTorch, Keras and TensorFlow dominate others with lots of variation in between.
+Bayesian and deep learning approaches are related but very different technologies in goals, implementation and applicability with little actual overlap so this is not a surprise, e.g., deep learning cannot bring the explainability of applied math/statistics and Bayesian methods do not scale to deep learning data sets.
+
+While deep learning behemoths like Facebook and Google both use and support Bayesian efforts, the Bayesian packages scientists actually use are academic/volunteer efforts punching far above their weight class and they need financial support.
+It would behoove funders to fully understand the impact and role of Bayesian methods in resource allocation.
+-->
+
 # Introduction
@@
 This is not in our collective imagination--Bayesians like to 'believe things' after all--NSF reviews came back dismissing Stan funding because all the interesting work was assumed to be happening with deep learning.
-Recently in the UK, open skepticism was expressed about impact claims for Bayesian software in response to a research grant.
+Recently in the UK, open skepticism was expressed about impact claims for Bayesian software in response to a research grant[^0].
 Mind you, deep learning and Bayesian modeling are conceptual cousins but in the end are very different from each other.
@@
 Yet Bayesians found themselves in deep learning's shadow somehow.
+
+[^0]: <!--R0: -->I will avoid substantiating the rejection of grants due to the dominance of deep learning with the names of my sources.
+I am not comfortable revealing private conversations that are perhaps best kept in confidence.
+In my own grant writing efforts I have not seen such comments but I have defensively addressed the deep learning issues due to what had been communicated to me.
+<!--:R0-->
 
@@
 This analysis led me to revise that opinion considerably.
+
+<!--R4:-->
+
+## One of these things is not like the other: Characterizing the difference between Bayesian modeling and deep learning
+
+What, if anything, are the hard differences between Bayesian modeling and deep learning?
+Deep learning can be implemented in Bayesian models, and lots of Bayesian concepts get used in the deep learning world as one reviewer noted:
+
+"There is a *huge* rise in the use of Bayesian Deep learning.
+Please do not fall into the trap of making DL look like some ‘other’ thing.
+It’s just a non/hyper-parametric model and as such we can do all the full Bayesian stuff - and many do!
+There are well-documented works on semi-structured HMC, Riemann MC and good old fashioned MCMC for use in big deep nets, let alone all the neat work on approximate Bayes, mainly using variational learning.
+Look at work on things like loss-calibrated Bayesian deep nets, Bayes DL, Bayesian autoencoding, ‘HamilTorch’ MCMC package for TensorFlow (https://adamcobb.github.io/journal/hamiltorch.html) etc."
+
+Perhaps Stan and PyMC3 are really part of the same discipline as deep learning and this article catalogs usage differences of siblings from the same parents--sort of like the difference between R and Python?
+
+But no.
+Another reviewer comment contains the key insight: they said it was no surprise that Bayesian software had high usage outside of computer science since other subjects are "dominated by Stats & Applied Math".
+I took this to mean that deep learning is not a natural candidate for use in those fields, and software like Stan and PyMC3 are.
+Unpacking that a bit I'll observe:
+
+* Unless the phenomenon under study is a physical neural net, deep learning only offers prediction services.
+While outstanding progress has been made and further progress may well involve Bayesian concepts, it is prediction in the end.
+* The basis of prediction in deep learning is opaque to human comprehension even if Bayesian techniques are used. This applies to generative neural nets as well. 
+* Opacity blocks use in fields "dominated by Stats & Applied Math" where the goal involves developing and fitting mechanistic models for the most part.
+The science is in the model description, the quality of fit validates the model.
+A high quality fit in the absence of an understandable model does not help.
+
+Deep learning clearly is used in mechanistic models but typically as a sensor or classifier, e.g., classify x-ray images for evidence of COVID pneumonia in an epidemiological study.
+That study itself will likely be a state-based model where transition rates are explicitly estimated statistically and are human interpretable.
+
+Another case are hybrid models where a deep learning component replaces some or all of the likelihood of a Bayesian model.
+While the overall model may be interpretable, the individual deep learning components remain black boxes.
+Interpretability for the overall model comes from have well understood properties of the deep learning components, e.g., how they were trained etc.
+For example Maggie Lieu's presentation at StanCon that used a deep learning component to substitute for a numerically unstable halo mass function (@lieuAstro).
+
+Also Bayesian models can be utterly opaque if not authored with an eye to understandability, but the option for understandability exists and is generally the expectation.
+Deep learning systems cannot match this because a big portion of 'authoring' (hyper parameter setting) in those frameworks are beyond human understanding.
+
+In the end I believe model interpretability cleanly differentiates the the deep learners and Bayesian modelers[^3], at least in practice.
+
+[^3]:However, in a year this article may be much more difficult to write since deep learning packages have adopted HMC implementations.
+But I'd suggest that is more of a grafting of another approach to their code base rather than natural growth of deep learning techniques into the PyMC3/Stan space.
+
+<!--:R4-->
+
+## Defining what counts as a citation
 
@@
 The increased counts are likely due to inclusion of non-peer reviewed sites like https://arxiv.org
-
-<br>
 
@@
 
+<!--R1:--> This paper also demonstrates that deep learning is not the default tool when it is time for some reasoning in the scientific enterprise.
+If you are doing immunology/microbiology then Bayesian methods dominate.
+Environmental science, economics are a coin flip.
+While the choice of machine reasoning techniques should be entirely determined by the problem being addressed, researchers use what is popular and being talked about.
+Perhaps this paper gets Bayesian methods considered a bit more.
+<!--:R1-->
+
 As a Bayesian advocate I'd like to leave the funding agencies with the observation that Bayesian software carries roughly one third of the research load by research citation count (3925/(3925+8884)) without the backing of Fortune 500 companies like Facebook and Google.
@@
 Supporting technology for this document includes the R language (@Rcite), the 'tidyverse' group of packages (@tidyverse) that give data and visualization tools and this Rmarkdown document (@rmarkdown) rendered with 'bookdown' (@bookdown) developed using the Rstudio (@rstudio) IDE (Integrated Development Environment).
-Supporting package include 'kable-extra' (@kable-extra), a table formatter, the 'scales' (@scales) package for converting to percentages, 'ggrepel' (@ggrepel) for labeling line graphs with dynamically shifting labels, 'jsonlite' (@jsonlite) for JSON data parsing, 'httr' (@httr) for GET requests, 'stringr' (@stringr) for regular expression matching and 'redux' (@redux) for access to the Redis server (@redis) which I highly recomend using as a caching layer for this sort of project.
+Supporting package include 'kable-extra' (@kable-extra), a table formatter, the 'scales' (@scales) package for converting to percentages, 'ggrepel' (@ggrepel) for labeling line graphs with dynamically shifting labels, 'jsonlite' (@jsonlite) for JSON data parsing, 'httr' (@httr) for GET requests, 'stringr' (@stringr) for regular expression matching and 'redux' (@redux) for access to the Redis server (@redis) which I highly recommend using as a caching layer for this sort of project.
 
@@
 Oriol Abril Pla brought up emcee as a popular Bayesian system worth consideration which was a big oversight on my part as well as providing insightful commentary.
+<!--R5-->: I'd also like to thank anonymous reviewers who substantially improved the exposition.
+<!--:R5-->
 
 I intend this to be an evolving document with periodic updates at it's github repo at https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html.
-Note that github pages changes urls around in odd ways for the rendered view of hte html page.
+Note that github pages changes urls around in odd ways for the rendered view of the html page.
 All source code is viewable in the .Rmd document.

### 2021-11-15 DeepLearningAndBayesianSoftware.Rmd
try 2 at final

@@
 subtitle: "Comparing research use via citation counting"
-author: "Breck Baldwin"
+author: "Breck Baldwin, Columbia University"
 email: "breckbaldwin@gmail.com"
-date: "11/01/2021"
+affiliation: Columbia Univeristy
+
 output:
@@
 ```
+
+<!--
+
+    A short informative title containing the major key words. The title should not contain abbreviations (see Wiley's best practice SEO tips);
+    A short running title of less than 40 characters;
+    The full names of the authors;
+    The author's institutional affiliations where the work was conducted, with a footnote for the author’s present address if different from where the work was conducted;
+    Acknowledgements;
+    Abstract and keywords;
+    Main text;
+    References;
+    Tables (each table complete with title and footnotes);
+    Figure legends;
+    Appendices (if relevant).
+-->
 
@@
 
-Another case are hybrid models where a deep learning component replaces some or all of the likelihood of a Bayesian model.
+Another case includes hybrid models where a deep learning component replaces some or all of the likelihood of a Bayesian model.
 While the overall model may be interpretable, the individual deep learning components remain black boxes.
-Interpretability for the overall model comes from have well understood properties of the deep learning components, e.g., how they were trained etc.
-For example Maggie Lieu's presentation at StanCon that used a deep learning component to substitute for a numerically unstable halo mass function (@lieuAstro).
+Interpretability for the overall model comes via having well understood properties for the deep learning components, e.g., how they were trained etc.
+or other methods of characterizing their role.
+For example Maggie Lieu's presentation at StanCon used a deep learning component to substitute for a numerically unstable halo mass function (@lieuAstro).
 
@@
 
-In the end I believe model interpretability cleanly differentiates the the deep learners and Bayesian modelers[^3], at least in practice.
+In the end I believe model interpretability cleanly differentiates the deep learners and Bayesian modelers[^3], at least in practice.
 
@@
 Since Bayesian modeling seems always in the shadow of deep learning I started tracking deep learning software packages as well for comparison.
-Below we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with support/derivative packages RStan, RStanArm, PyStan and brms.
+Below we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with supporting/derivative packages RStan, RStanArm, PyStan and brms.
 
@@
   scale_color_discrete(guide = FALSE) +
-  labs(y = "year count for research articles matching topic",
+  labs(y = "Year count for research articles matching topic",
   caption = "Fig 1: Relative counts with and without computer science journals") +
@@
                      trans = scales::log_trans())
+ggsave('fig1.png', plot=plot2)
 print(plot2)
@@
 Direct citation linking does not work well with these packages since many of them had no journal publication to cite in a referring article.
-As a result a string search was used to identify the citations in the reference section which was determined by Scopus.com's document parser.
+Consequently a string search was used to identify the citations in the reference section which was determined by Scopus.com's document parser.
 I would estimate a 5% false positive rate based based on informal examination of return sets.
@@
 
-The above queries can be run with the advance search option at https://scopus.com and are generated from the API queries actually used where '+' was replaced with ' ' above.
+The above queries can be run with the advance search option at https://scopus.com and are generated from the API queries actually used where '+' is replaced with ' ' above.
 An account is required which is included for many education institutions.
@@
 
-# Access to source/scripts and acknowledgements
+# Access to source/scripts
 
@@
 
-I'd like to thank Andrew Gelman and the Laplace lab at Columbia University for support.
+I intend this to be an evolving document with periodic updates at it's github repo at https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html.
+Note that github pages changes urls around in odd ways for the rendered view of the html page.
+All source code is viewable in the .Rmd document.
+
+# Acknowledgements
+
+I'd like to thank Andrew Gelman and the Laplace lab at Columbia University. The research was supported by internal Columbia funds.
 
@@
 
-I intend this to be an evolving document with periodic updates at it's github repo at https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html.
-Note that github pages changes urls around in odd ways for the rendered view of the html page.
-All source code is viewable in the .Rmd document.
-
 I can be reached at breckbaldwin@gmail.com or via the github issues on the repository.

### 2021-12-23 DeepLearningAndBayesianSoftware.Rmd
stand alone version of paper

@@
 
-USE_CACHE = TRUE #will have to setup a redis server, strongly recommend doing this
+USE_CACHE = FALSE #will have to setup a redis server, strongly recommend doing this
 REPORT_PROGRESS = FALSE
+RUN_QUERY <- FALSE #if TRUE will either hit scopus or cache, if false will pull from disk
 
@@
 
-```{r DL vs Bayes Graph, echo=FALSE, message=FALSE, warning=FALSE}
+```{r DL vs Bayes Data Creation, echo=FALSE, message=FALSE, warning=FALSE}
 library(ggrepel)
@@
 years <- year_start:year_end
+
+if (RUN_QUERY) {
   scopus.df <- data.frame(years)
@@
 
+  write.table(scopus.df,"scopus.csv")
+}
+
+scopus.df <- NA
+scopus.df <- read.table("scopus.csv")
 column_names <- colnames(scopus.df)
@@
 The next question is what does the the distribution look like for non-computer science subject areas?
-Below is a table of subject counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed:
+Below in Table 1 are subject category counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed:
 
@@
 
+if (RUN_QUERY) {
   bayesVdeepLearnDf <- getAllSubjectCategories(queryBayVsDeepM)
@@
   allDataDf <- allDataDf[,c(1,2,3,11,6,7,10,9,4,8,5,12,13)]
+  write.table(allDataDf,"allDataDf.csv")
+}
+
+allDataDf <- read.table("allDataDf.csv")
 displayedColNames <- colnames(allDataDf)

### 2021-12-26 Submission.tex
version added to get past journal of AI typesetters

--- 2021-11-15 DeepLearningAndBayesianSoftware.Rmd
@@
----
-title: "Deep Learning does not Replace Bayesian Modeling"
-subtitle: "Comparing research use via citation counting"
-author: "Breck Baldwin, Columbia University"
-email: "breckbaldwin@gmail.com"
-affiliation: Columbia Univeristy
+\documentclass[AMA,STIX1COL]{WileyNJD-v2}
 
-output:
- bookdown::html_document2:
-   includes:
-     in_header: _html/ga.html
+\articletype{Article Type}%
+%\usepackage{dirtytalk}
+\usepackage{graphicx}
+\received{}
+\revised{}
+\accepted{}
 
-bibliography: citations.bib
-csl: american-medical-association.csl
-link-citations: yes
----
+\raggedbottom
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(
-	echo = TRUE,
-	message = FALSE,
-	comment = NA
-)
-```
+\begin{document}
 
-<!--
+\title{Deep Learning does not Replace Bayesian Modeling: Comparing research use via citation counting}
 
-    A short informative title containing the major key words. The title should not contain abbreviations (see Wiley's best practice SEO tips);
-    A short running title of less than 40 characters;
-    The full names of the authors;
-    The author's institutional affiliations where the work was conducted, with a footnote for the author’s present address if different from where the work was conducted;
-    Acknowledgements;
-    Abstract and keywords;
-    Main text;
-    References;
-    Tables (each table complete with title and footnotes);
-    Figure legends;
-    Appendices (if relevant).
--->
+\author[1]{Breck Baldwin*}
 
-<!--**Abstract revised: Note this is not in the paper but a separate attachment**
+% \authormark{AUTHOR ONE \textsc{et al}}
 
-One could be excused for assuming that deep learning had or will soon usurp all credible work in reasoning, artificial intelligence and statistics, but like most 'meme' class broad generalizations the concept does not hold up to scrutiny.
+\address[1]{\orgname{Columbia University}, \orgaddress{\state{New York}, \country{USA}}}
+
+\corres{*Breck Baldwin, \email{breckbaldwin@gmail.com}}
+
+\presentaddress{Safety 3rd}
+
+\abstract[Summary]{One could be excused for assuming that deep learning had or will soon usurp all credible work in reasoning, artificial intelligence and statistics, but like most 'meme' class broad generalizations the concept does not hold up to scrutiny.
 Memes don't generally matter since the experts will always know better but in the case of Bayesian software like Stan and PyMC3 even its developers and advocates bemoan the apparent dominance of deep learning as manifested in popular culture, breathtaking performance and most problematically from funding agency peer review that impacts our ability to further advance the field.
@@
 While deep learning behemoths like Facebook and Google both use and support Bayesian efforts, the Bayesian packages scientists actually use are academic/volunteer efforts punching far above their weight class and they need financial support.
-It would behoove funders to fully understand the impact and role of Bayesian methods in resource allocation.
--->
+It would behoove funders to fully understand the impact and role of Bayesian methods in resource allocation.}
 
-# Introduction
+\keywords{Bayesian, Stan, PyMC3, Deep Learning}
+
+\maketitle
+
+\section{Introduction}\label{sec1}
 
 Stan, a Bayesian modeling language, was released in 2012 to considerable fanfare.
-A shiny new inference algorithm, HMC with NUTS (@hoffman2014no), promised and delivered fits that could not be fit before.
+A shiny new inference algorithm, HMC (Hameltonian Monte-Carlo) with NUTS (No U-Turn Sampler) \cite{hoffman2014no}, promised and delivered fits that could not be fit before.
 And then deep learning happened.
-The release of TensorFlow (@abadi2016tensorflow) opened the doors of deep learning to all and by it 2016 began getting significant traction.
+The release of TensorFlow \cite{abadi2016tensorflow} opened the doors of deep learning to all and by it 2016 began getting significant traction.
 One result of this was people often assumed that deep learning's successes usurped Bayesian modeling's domain.
 This is not in our collective imagination--Bayesians like to 'believe things' after all--NSF reviews came back dismissing Stan funding because all the interesting work was assumed to be happening with deep learning.
-Recently in the UK, open skepticism was expressed about impact claims for Bayesian software in response to a research grant[^0].
+Recently in the UK, open skepticism was expressed about impact claims for Bayesian software in response to a research grant \footnote{\label{one}I will avoid substantiating the rejection of grants due to the dominance of deep learning with the names of my sources.
+I am not comfortable revealing private conversations that are perhaps best kept in confidence.
+In my own grant writing efforts I have not seen such comments but I have defensively addressed the deep learning issues due to what had been communicated to me.}.
 Mind you, deep learning and Bayesian modeling are conceptual cousins but in the end are very different from each other.
@@
 
-[^0]: <!--R0: -->I will avoid substantiating the rejection of grants due to the dominance of deep learning with the names of my sources.
-I am not comfortable revealing private conversations that are perhaps best kept in confidence.
-In my own grant writing efforts I have not seen such comments but I have defensively addressed the deep learning issues due to what had been communicated to me.
-<!--:R0-->
-
 It is 2021 and this document does some very simple analysis around use of Bayesian and deep learning packages as evidenced in the research literature to get a perspective on what actually happened and is happening.
@@
 
-1.
-How does Bayesian modeling software stack up against deep learning without appeal to feature comparisons, performance arguments on suspect data sets or achieved closeness to Platonic ideals?
+\begin{enumerate}
+    \item How does Bayesian modeling software stack up against deep learning without appeal to feature comparisons, performance arguments on suspect data sets or achieved closeness to Platonic ideals?
 My response is--just go out and count how many citations the respective approaches have in the research literature.
 Counting and categorization, that's it.
-2. Asses the impact of Bayesian modeling software using deep learning metrics as a yard stick--how big a fraction of a huge thing are we?
-3. Contextualize the roles each approach has by looking at subject distributions. The technologies have very different use cases so one would expect variation.
+    \item Asses the impact of Bayesian modeling software using deep learning metrics as a yard stick--how big a fraction of a huge thing are we?
+    \item Contextualize the roles each approach has by looking at subject distributions. The technologies have very different use cases so one would expect variation.
+\end{enumerate}
 
@@
 
-<!--R4:-->
-
-## One of these things is not like the other: Characterizing the difference between Bayesian modeling and deep learning
+\subsection{One of these things is not like the other: Characterizing the difference between Bayesian modeling and deep learning}
 
@@
 
-"There is a *huge* rise in the use of Bayesian Deep learning.
+\begin{quote}
+There is a *huge* rise in the use of Bayesian Deep learning.
 Please do not fall into the trap of making DL look like some ‘other’ thing.
@@
 There are well-documented works on semi-structured HMC, Riemann MC and good old fashioned MCMC for use in big deep nets, let alone all the neat work on approximate Bayes, mainly using variational learning.
-Look at work on things like loss-calibrated Bayesian deep nets, Bayes DL, Bayesian autoencoding, ‘HamilTorch’ MCMC package for TensorFlow (https://adamcobb.github.io/journal/hamiltorch.html) etc."
+Look at work on things like loss-calibrated Bayesian deep nets, Bayes DL, Bayesian autoencoding, ‘HamilTorch’ MCMC package for TensorFlow (https://adamcobb.github.io/journal/hamiltorch.html) etc.
+\end{quote}
 
@@
 But no.
-Another reviewer comment contains the key insight: they said it was no surprise that Bayesian software had high usage outside of computer science since other subjects are "dominated by Stats & Applied Math".
+Another reviewer comment contains the key insight: they said it was no surprise that Bayesian software had high usage outside of computer science since other subjects are "dominated by Stats \& Applied Math".
 I took this to mean that deep learning is not a natural candidate for use in those fields, and software like Stan and PyMC3 are.
@@
 
-* Unless the phenomenon under study is a physical neural net, deep learning only offers prediction services.
+\begin{itemize}
+    \item Unless the phenomenon under study is a physical neural net, deep learning only offers prediction services.
 While outstanding progress has been made and further progress may well involve Bayesian concepts, it is prediction in the end.
-* The basis of prediction in deep learning is opaque to human comprehension even if Bayesian techniques are used. This applies to generative neural nets as well. 
-* Opacity blocks use in fields "dominated by Stats & Applied Math" where the goal involves developing and fitting mechanistic models for the most part.
+    \item The basis of prediction in deep learning is opaque to human comprehension even if Bayesian techniques are used. This applies to generative neural nets as well. 
+    \item Opacity blocks use in fields "dominated by Stats \& Applied Math" where the goal involves developing and fitting mechanistic models for the most part.
 The science is in the model description, the quality of fit validates the model.
 A high quality fit in the absence of an understandable model does not help.
+\end{itemize}
 
@@
 or other methods of characterizing their role.
-For example Maggie Lieu's presentation at StanCon used a deep learning component to substitute for a numerically unstable halo mass function (@lieuAstro).
+For example Maggie Lieu's presentation at StanCon used a deep learning component to substitute for a numerically unstable halo mass function \cite{lieuAstro}.
 
@@
 
-In the end I believe model interpretability cleanly differentiates the deep learners and Bayesian modelers[^3], at least in practice.
+In the end I believe model interpretability cleanly differentiates the deep learners and Bayesian modelers\footnote{\label{two}However, in a year this article may be much more difficult to write since deep learning packages have adopted HMC implementations.
+But I'd suggest that is more of a grafting of another approach to their code base rather than natural growth of deep learning techniques into the PyMC3/Stan space.}, at least in practice.
 
-[^3]:However, in a year this article may be much more difficult to write since deep learning packages have adopted HMC implementations.
-But I'd suggest that is more of a grafting of another approach to their code base rather than natural growth of deep learning techniques into the PyMC3/Stan space.
+\subsection{Defining what counts as a citation}
 
-<!--:R4-->
-
-## Defining what counts as a citation
-
-For the purposes of this document we take Bayesian software to be Stan (@standev) and PyMC3 (@salvatier2016probabilistic) with ecosystem components included that are likely to be cited.
-That includes the simplified syntax interfaces to Stan, brms (@brms), RstanArm (@rstanarm), and interface packages to Stan, RStan (@rstan) and PyStan (@pystan)[^1].
-The analysis applies only to packages that are in current development so the venerated, and very high impact, packages like BUGS (@lunn2013bugs) and JAGS (@plummer2003jags) are not considered although including them would double our counts.
-There is also the best named Bayesian package of all time, "emcee: The MCMC Hammer", (@foreman2013emcee) which enjoys tremendous use in the astrophysics community but is too specialized to be considered a general Bayesian package so it is not included.
+For the purposes of this document we take Bayesian software to be Stan \cite{standev} and PyMC3 \cite{salvatier2016probabilistic} with ecosystem components included that are likely to be cited.
+That includes the simplified syntax interfaces to Stan\footnote{\label{three}There are also lighter weight interfaces as well interface to other languages that I ignore due to low expected research mentions, see https://https://mc-stan.org/users/interfaces/ for a listing.}, brms \cite{brms}, RstanArm \cite{rstanarm}, and interface packages to Stan, RStan \cite{rstan} and PyStan \cite{pystan}.
+The analysis applies only to packages that are in current development so the venerated, and very high impact, packages like BUGS \cite{lunn2013bugs} and JAGS \cite{plummer2003jags} are not considered although including them would double our counts.
+There is also the best named Bayesian package of all time, "emcee: The MCMC Hammer", \cite{foreman2013emcee} which enjoys tremendous use in the astrophysics community but is too specialized to be considered a general Bayesian package so it is not included.
 Also, the ecosystems are actually much larger but they are unlikely to be cited in the research literature and we have to stop some place.
-For deep learning we take TensorFlow, its interface Keras (@keras) and PyTorch (@pytorch) as roughly equivalent entities for the comparison.
-Theano (@theano) is no longer in development.
+For deep learning we take TensorFlow, its interface Keras \cite{keras} and PyTorch \cite{pytorch} as roughly equivalent entities for the comparison.
+Theano \cite{theano} is no longer in development.
 It should be noted that both PyTorch and TensorFlow have implemented HMC with NUTS for Bayesian inference but those are recent developments that have not made much of an impact yet.
 
-[^1]:There are also lighter weight interfaces as well interface to other languages that I ignore due to low expected research mentions, see https://https://mc-stan.org/users/interfaces/ for a listing.
-
-The key resource behind this document is Elsevier's https://scopus.com research search engine that provides a tightly curated[^2] search index that includes sources outside of the academic behemoth's own journals.
+The key resource behind this document is Elsevier's https://scopus.com research search engine that provides a tightly curated\footnote{\label{four}https://scholar.google.com generally yields double the counts for similar queries but offers no subject classification or API to code against.
+The increased counts are likely due to inclusion of non-peer reviewed sites like https://arxiv.org} search index that includes sources outside of the academic behemoth's own journals.
 It also provides a solid API (Application Programmer Interface) and a classification of journals into subject areas.
@@
 
-[^2]:https://scolar.google.com genrally yields double the counts for similar queries but no subject classification or API to code against.
-The increased counts are likely due to inclusion of non-peer reviewed sites like https://arxiv.org
-
-# High level prevalence of Deep Learning vs Bayesian Modeling
+\section{High level prevalence of Deep Learning vs Bayesian Modeling}
 
@@
 Since Bayesian modeling seems always in the shadow of deep learning I started tracking deep learning software packages as well for comparison.
-Below we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with supporting/derivative packages RStan, RStanArm, PyStan and brms.
+In Figure 1 we see the relative citations of the top deep learning packages TensorFlow, PyTorch and the support package Keras to the Bayesian packages PyMC3, Stan with supporting/derivative packages RStan, RStanArm, PyStan and brms.
 
-```{r Access Scopus, include=FALSE}
-rm(list = rm()) #clean out any state from previous runs
-library(tidyverse)
-library(jsonlite)
-library(httr)
-library(kableExtra)
-library(stringr)
-
-# remember to set working directory to the location of this file
-setwd("~/git/ScientificSoftwareImpactMetrics")
-credentials <- read_json('Scopus_credentials.json')
-API_KEY <- credentials$API_KEY
-INSTITUTION_TOKEN <- credentials$INSTITUTION_TOKEN
-# Format of Scopus_credentials.json
-# {
-# "API_KEY":"XXXXXXXXXXXXXXXXXXXXXXXXXxx",
-# "INSTITUTION_TOKEN":"XXXXXXXXXXXXXXXXXXXXXXXX"
-# }
-
-BASE_URL = 'https://api.elsevier.com/content/search/scopus'
-
-USE_CACHE = TRUE #will have to setup a redis server, strongly recommend doing this
-REPORT_PROGRESS = FALSE
-
-if (USE_CACHE) {
-  library(redux)
-  redis <- redux::hiredis()
-  get_results <- function(url) {
-    cache <- redis$GET(url) # check redis first
-    if (!is.null(cache)) {
-      result <- unserialize(cache)
-      if (result$status_code == 200) {
-        if (REPORT_PROGRESS) {
-          cat(paste("\nhitting redis for", url))
-        }
-        return(result)
-      }
-    }
-    random_wait <- abs(rnorm(1, 1, 1))
-    if (REPORT_PROGRESS) {
-      cat(paste("\ncache miss, querying:", url, "\n"))
-      cat(paste(
-        "\nWaiting",
-        random_wait,
-        "seconds to be nice to webserver\n"
-      ))
-    }
-    Sys.sleep(random_wait)
-    result <- GET(
-      url,
-      add_headers('X-ELS-APIKey' = API_KEY,
-                  'X-ELS-Insttoken' = INSTITUTION_TOKEN)
-    )
-    if (result$status_code != 200) { 
-      stop(sprintf("Got error from search, got status code %d",result$status_code))
-    }
-    redis$SET(url, serialize(result, NULL))
-    return(result)
-  }
-} else { #not using cache, hacky but works with R
-  get_results <- function(url) {
-    random_wait <- abs(rnorm(1, 1, 1))
-    if (REPORT_PROGRESS) {
-      cat(paste("\nno cache used, querying:", url, "\n"))
-      cat(paste(
-        "\nWaiting",
-        random_wait,
-        "seconds to be nice to webserver\n"
-      ))
-    }
-    Sys.sleep(random_wait)
-    result <- GET(
-      url,
-      add_headers('X-ELS-APIKey' = API_KEY,
-                  'X-ELS-Insttoken' = INSTITUTION_TOKEN)
-    )
-    if (result$status_code != 200) { 
-      stop(sprintf("Got error from search, got status code %d",result$status_code))
-    }
-    return(result)
-  }
-}
-
-# setup queries for analysis
-
-stan_q <- '(gelman+AND+hoffman+AND+stan)+OR+mc-stan.org'
-pymc_q <- 'PyMC3+OR+(PyMC*+AND+fonnesbeck)'
-rstan_q <- 'rstan+AND+NOT+mit'
-brms_q <-  'brms+AND+burkner'
-emcee_q <- 'emcee+AND+foreman-mackey'
-
-#Display names followed by queries
-pkg_query <- c('emcee', 'Stan', 'brms','PyStan', 'RStanArm', 'RStan','PyMC',  'TensorFlow', 'PyTorch', 'Keras',
-               emcee_q, stan_q, brms_q, 'pystan', 'rstanarm', rstan_q, pymc_q, 'tensorflow', 'pytorch', 'Keras')
-pkg_query_m <- matrix(data = pkg_query[c(-1,-11)], ncol =2) #emcee removed
-queryBayVsDeepM <- matrix(ncol = 2, nrow = 2)
-queryBayVsDeepM[1,] <- c('Bayesian', 
-                         paste0('(', paste(pkg_query_m[1:6,2], collapse = ")+OR+("),')'))
-queryBayVsDeepM[2,] <- c('Deep Learning', 
-                         paste0('(', paste(pkg_query_m[7:9,2], collapse = ")+OR+("), ')'))
-
-```
-
-```{r DL vs Bayes Graph, echo=FALSE, message=FALSE, warning=FALSE}
-library(ggrepel)
-year_start <- 2012
-year_end <- 2020
-years <- year_start:year_end
-df <- data.frame(years)
-
-years <- year_start:year_end
-scopus.df <- data.frame(years)
-#queryM = pkg_query_m
-queryM = queryBayVsDeepM
-for (i in 1:nrow(queryM)) {
-  package_name = queryM[i, 1]
-  query = queryM[i, 2]
-  for (suffix in c('','_no_comp_sci')) {
-#  for (suffix in c('')) {
-    year_counts = c()
-    for (year in year_start:year_end) {
-      url <- paste(BASE_URL, '?query=','REF(', query,')',
-                   "+AND+PUBYEAR+=+", year,
-                   sep = '')
-      if (suffix == '_no_comp_sci') {
-        url = paste(url, '+AND+NOT+SUBJAREA(COMP)', sep = '')
-      }
-      result <- get_results(url)
-      json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))
-      data <- jsonlite::fromJSON(json_txt)
-      year_counts = c(
-        year_counts,
-        as.numeric(data$`search-results`$`opensearch:totalResults`))
-    }
-    scopus.df[paste(
-      package_name,
-      suffix,
-      sep = '')] <- year_counts
-  }
-}
-
-column_names <- colnames(scopus.df)
-col_sums <- colSums(scopus.df)
-df_long <- gather(scopus.df, topic, yr_count,
-                  column_names[2]:column_names[length(column_names)])
-
-df_long_label <- df_long %>%
-  mutate(label = if_else(years ==  max(years),
-                         paste(as.character(topic),col_sums[topic], sep = ':'), NA_character_))
-
-plot2 <-
-  ggplot(data = df_long_label, aes(
-    x = years,
-    y = yr_count,
-    group = topic,
-    color = topic
-  )) +
-  geom_line() +
-  geom_point() +
-  geom_label_repel(aes(label = label),
-                   na.rm = TRUE) +
-  scale_color_discrete(guide = FALSE) +
-  labs(y = "Year count for research articles matching topic",
-  caption = "Fig 1: Relative counts with and without computer science journals") +
-   theme(plot.caption=element_text(size=12, hjust=0, margin=margin(15,0,0,0)))
-
-   log_plot = plot2 + 
-     scale_y_continuous(breaks=c(1, 10, 100, 1000, 10000, 100000, 1e+06, 1e+07, 1e+08), 
-                     trans = scales::log_trans())
-ggsave('fig1.png', plot=plot2)
-print(plot2)
-```
-
-Reading from top of Fig 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,600--nearly a 10x difference.
+Reading from top of Figure 1, assuming raw research citations counts matter, Bayesian packages are overwhelmed by deep learning packages with a count of nearly 35,000 citations to a Bayesian count of 4,600--nearly a 10x difference.
 One could excuse a bit of Bayesian despondency in the face of the wave depicted above--this ratio supports my expectations.
-But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 35k to 8.8k--a 75% drop.
+But graph also shows relative counts for research citations outside of computer science which drops the count of deep learning from 35k to 8.8k--a 75\% drop.
 That drop IS a surprise.
 I didn't have a prior on deep learning's distribution across subject categories but I didn't expect it to be so skewed.
-The Bayesian ecosystem shows a total of 4.5k of which 14% are computer science--that ratio does not surprise me.
+The Bayesian ecosystem shows a total of 4.5k of which 14\% are computer science--that ratio does not surprise me.
 
-# Subject Category Breakdowns
+\section{Subject Category Breakdowns}
 
@@
 The next question is what does the the distribution look like for non-computer science subject areas?
-Below is a table of subject counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed:
+Below in Table 1 are subject category counts for packages starting in 2006 when PyMC was first released to 2021 with relative percentages listed.
 
-```{r Table 1, echo=FALSE, message=FALSE, warning=FALSE}
-library(scales)
-library(kableExtra)
-
-makeDisplayTable <- function(df) {
-  subjCats = colnames(df)[-1:-2]
-  df %>% group_by(package) %>% 
-    summarize_each(funs(sum), all_of(subjCats)) -> outDf
-  finalDf = t(outDf)
-  subFrame = finalDf[-1,]
-  cleanDf =  as.data.frame(apply(subFrame, 2, as.numeric))
-  colnames(cleanDf) = finalDf[1,]
-  rownames(cleanDf) = rownames(finalDf)[-1]
-  cleanDf %>% mutate(totals = rowSums(.)) -> summedDf
-  displayDf = summedDf
-  displayDf[] = NA_character_
-  for (i in 1:nrow(summedDf)) {
-    displayDf[i,] = as.list(sprintf("%d/%s", as.numeric(summedDf[i,]), 
-                                    label_percent(accuracy = 1)(as.numeric(summedDf[i,]/summedDf[i,]$totals))))
-    displayDf[i,]$totals = summedDf[i,]$totals 
-  }
-  displayOrderedDf <- displayDf[order(row.names(displayDf)),] 
-  return(displayOrderedDf)
-}
-
-getAllSubjectCategories <- function (queryM) {
-  year_start = 2006
-  year_end = 2021
-  years <- year_start:year_end
-  accumDf = data.frame(matrix(ncol = 2, nrow = nrow(queryM) * length(years)))
-  colnames(accumDf) = c('package', 'years')
-  counter = 1
-  article_count = 0
-  article_subj_count = 0
-  for (i in 1:nrow(queryM)) {
-    package_name = queryM[i, 1]
-    query = queryM[i, 2]
-    accumDf[counter:((counter - 1) + length(years)),]$years <- years
-    accumDf[counter:((counter - 1) + length(years)),]$package <- 
-      rep(package_name, length(years))
-    for (year in year_start:year_end) {
-      url <-
-        paste(
-          BASE_URL,
-          '?query=',
-          'REF(',
-          query,
-          ')',
-          "+AND+PUBYEAR+=+",
-          year,
-          '&facets=subjarea(count=101)',
-          sep = ''
-        )
-      result <- get_results(url)
-      json_txt <- rawToChar(as.raw(strtoi(result$content, 16L)))
-      data <- jsonlite::fromJSON(json_txt)
-      article_count <-
-        as.numeric(data$`search-results`$`opensearch:totalResults`) + article_count
-      facet_count <- length(data$`search-results`$facet$category$name)
-      j <- 1
-      while (j < facet_count) {
-        name <- data$`search-results`$facet$category$label[j]
-        name <- str_replace(name, " \\(all\\)", "")
-        hitCount <- as.numeric(data$`search-results`$facet$category$hitCount[j])
-        article_subj_count <- hitCount + article_subj_count
-        if (!name %in% colnames(accumDf)) {
-          accumDf[name] <- rep(0, length(years) * nrow(queryM))
-          #print(paste("name=",name,", count=",hitCount))
-        }
-        accumDf[counter,][[name]] <- hitCount
-        j <- j+ 1
-      }
-      counter = counter + 1
-    }
-  }
-  if (REPORT_PROGRESS) {
-    print(sprintf("%d paperXsubject with %d total papers for %.3f avg subjects per paper",
-                  article_subj_count,
-                  article_count,
-                  article_subj_count/article_count))
-  }
-  return(accumDf)
-}
-
-bayesVdeepLearnDf <- getAllSubjectCategories(queryBayVsDeepM)
-displayBayesVdeepLearnDf <- makeDisplayTable(bayesVdeepLearnDf)
-
-detailDf  <- getAllSubjectCategories(pkg_query_m)
-displayDetailDf <- makeDisplayTable(detailDf)
-
-allDataDf <- cbind(displayBayesVdeepLearnDf,displayDetailDf)
-allDataDf <- allDataDf[,c(1,2,3,11,6,7,10,9,4,8,5,12,13)]
-displayedColNames <- colnames(allDataDf)
-colnames(allDataDf) <- c(displayedColNames[1:12],'detail totals')
-kable(allDataDf, booktabs = TRUE) %>% kable_styling(font_size = 10)
-
-```
-
-Table 1 shows the basic subject split between Bayesian modeling and deep learning with total for the two categories, following are the individual package counts with a 'detail totals' as well for the individual packages.
-Note that the totals do not match because an article can match more than one package and as result gets contributes to the count of multiple packages.
-On average an article's journal is classified into two subject categories.
-
-Staring with the subject that got us here we see that computer science has 727 Bayesian publications vs 29,365 deep learning publications for 98% of the share of 30,092 total.
+Staring with the subject that got us here we see that computer science has 727 Bayesian publications vs 29,365 deep learning publications for 98\% of the share of 30,092 total.
 The left most two columns are the same queries used for the graph in Fig 1 with the total in column 3.
 Columns 4-12 are individual packages with percentage of the total in the 13th column.
-Sticking with computer science, one sees that TensorFlow has 18,178 references with 51% of the count--note that the total counts are higher because an article can matched by more that one query, each package gets a query, so the article count is 3 if Keras, PyMC and TensorFlow queries match the references and it has a computer science subject classification.
+Sticking with computer science, one sees that TensorFlow has 18,178 references with 51\% of the count--note that the total counts are higher because an article can matched by more that one query, each package gets a query, so the article count is 3 if Keras, PyMC and TensorFlow queries match the references and it has a computer science subject classification.
 
@@
 
-# Details of the queries
+\subsection{Details of the queries}
 
-Below are the queries used to identify the packages.
+In Table 2 are the queries used to identify the packages.
 Direct citation linking does not work well with these packages since many of them had no journal publication to cite in a referring article.
 Consequently a string search was used to identify the citations in the reference section which was determined by Scopus.com's document parser.
-I would estimate a 5% false positive rate based based on informal examination of return sets.
+I would estimate a 5\% false positive rate based based on informal examination of return sets.
 
-For an example of how the queries were developed, the stand alone query 'rstan' had the addition of rejecting articles that mentioned 'mit' because there was a common reference to 'http://wwwmath.mit.edu/~rstan/ec/' that existed prior to 2012 when Stan was released.
+For an example of how the queries were developed, the stand alone query 'rstan' clearly over-generated results due to articles that mentioned 'mit'.
+There is a common reference to 'http://wwwmath.mit.edu/\textasciitilde{}rstan/ec/' that existed prior to 2012 when Stan was released so the term 'mit' is negated in the resulting query.
 The name 'Stan' presented obvious difficulties, see https://statmodeling.stat.columbia.edu/2019/04/29/we-shouldntve-called-it-stan-i-shouldve-listened-to-bob-and-hadley/.
@@
 How the word 'mc-stan.org' was tokenized is unknown but there were no results before 2013 for the solo query.
-
-```{r Query Table, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
-pkg_query_m_display <- rbind(pkg_query_m, queryBayVsDeepM)
-pkg_query_m_display[,2] <- sprintf("REF(%s)", str_replace_all(pkg_query_m_display[,2], '\\+',' '))
-
-kable(pkg_query_m_display, booktabs = TRUE) %>% kable_styling(font_size = 12)
-
-```
 
@@
 
-# Conclusions
+\section{Conclusions}
 
-As often happens conducting a closer look reveals subtleties.
+As often happens, conducting a closer look reveals subtleties.
 Deep learning dominates the computer science part of science and obviously it dominates in the commercial sector which this article does not address.
@@
 
-<!--R1:--> This paper also demonstrates that deep learning is not the default tool when it is time for some reasoning in the scientific enterprise.
+This paper also demonstrates that deep learning is not the default tool when it is time for some reasoning in the scientific enterprise.
 If you are doing immunology/microbiology then Bayesian methods dominate.
@@
 Perhaps this paper gets Bayesian methods considered a bit more.
-<!--:R1-->
 
@@
 
-# Access to source/scripts
+\section{Access to source/scripts}
 
-Supporting technology for this document includes the R language (@Rcite), the 'tidyverse' group of packages (@tidyverse) that give data and visualization tools and this Rmarkdown document (@rmarkdown) rendered with 'bookdown' (@bookdown) developed using the Rstudio (@rstudio) IDE (Integrated Development Environment).
-Supporting package include 'kable-extra' (@kable-extra), a table formatter, the 'scales' (@scales) package for converting to percentages, 'ggrepel' (@ggrepel) for labeling line graphs with dynamically shifting labels, 'jsonlite' (@jsonlite) for JSON data parsing, 'httr' (@httr) for GET requests, 'stringr' (@stringr) for regular expression matching and 'redux' (@redux) for access to the Redis server (@redis) which I highly recommend using as a caching layer for this sort of project.
+Supporting technology for this document includes the R language \cite{Rcite}, the 'tidyverse' group of packages \cite{tidyverse} that give data and visualization tools and this Rmarkdown document \cite{rmarkdown} rendered with 'bookdown' \cite{bookdown} developed using the Rstudio \cite{rstudio} IDE (Integrated Development Environment).
+Supporting package include 'kable-extra' \cite{kable-extra}, a table formatter, the 'scales' \cite{scales} package for converting to percentages, 'ggrepel' \cite{ggrepel} for labeling line graphs with dynamically shifting labels, 'jsonlite' \cite{jsonlite} for JSON data parsing, 'httr' \cite{httr} for GET requests, 'stringr' \cite{stringr} for regular expression matching and 'redux' \cite{redux} for access to the Redis server \cite{redis} which I highly recommend using as a caching layer for this sort of project.
 
-I intend this to be an evolving document with periodic updates at it's github repo at https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html.
-Note that github pages changes urls around in odd ways for the rendered view of the html page.
+I intend this to be an evolving document with periodic updates at it's github repo \footnotetext{ https://github.com/breckbaldwin/ScientificSoftwareImpactMetrics where it lives as `DeepLearningAndBayesianSoftware.Rmd`, is rendered as `DeepLearningAndBayesianSoftware.html` and is viewable as https://breckbaldwin.github.io/ScientificSoftwareImpactMetrics/DeepLearningAndBayesianSoftware.html.} Note that github pages changes urls around in odd ways for the rendered view of the html page.
 All source code is viewable in the .Rmd document.
 
-# Acknowledgements
+\section{Acknowledgements}
 
@@
 Oriol Abril Pla brought up emcee as a popular Bayesian system worth consideration which was a big oversight on my part as well as providing insightful commentary.
-<!--R5-->: I'd also like to thank anonymous reviewers who substantially improved the exposition.
-<!--:R5-->
+I'd also like to thank anonymous reviewers who substantially improved the exposition.
 
@@
 
-# References
+\nocite{*}% Show all bib entries - both cited and uncited; comment this line to view only cited bib entries;
+\bibliography{citations}%
+
+\begin{table}
+\centering
+\caption{Subject Categories by packages\label{tab1}}
+\resizebox{\textwidth}{!}
+{\begin{tabular}{p{50}|l|p{40}|r||l|l|l|l|l|l|l|l|l|r}
+\hline
+  & Bayesian & Deep Learning & total & Stan & PyMC & PyStan & RStanArm & RStan & brms & PyTorch & Keras & TensorFlow & total\\
+\hline
+Computer Science & 727/2\% & 29365/98\% & 30092 & 463/1\% & 233/1\% & 15/0\% & 22/0\% & 76/0\% & 100/0\% & 7420/21\% & 9058/25\% & 18178/51\% & 35565\\
+\hline
+Agricultural and Biological Sciences & 1275/61\% & 820/39\% & 2095 & 821/29\% & 94/3\% & 3/0\% & 114/4\% & 365/13\% & 440/15\% & 102/4\% & 397/14\% & 534/19\% & 2870\\
+\hline
+Arts and Humanities & 385/23\% & 1305/77\% & 1690 & 209/10\% & 17/1\% & 1/0\% & 28/1\% & 78/4\% & 204/10\% & 389/19\% & 365/18\% & 729/36\% & 2020\\
+\hline
+Biochemistry, Genetics and Molecular Biology & 752/25\% & 2232/75\% & 2984 & 451/12\% & 133/4\% & 12/0\% & 63/2\% & 120/3\% & 213/6\% & 407/11\% & 951/25\% & 1424/38\% & 3774\\
+\hline
+Business, Management and Accounting & 92/13\% & 598/87\% & 690 & 59/7\% & 11/1\% & 0/0\% & 7/1\% & 20/2\% & 21/3\% & 107/13\% & 216/26\% & 397/47\% & 838\\
+\hline
+Chemical Engineering & 51/5\% & 933/95\% & 984 & 19/2\% & 27/2\% & 0/0\% & 1/0\% & 5/0\% & 3/0\% & 184/15\% & 383/31\% & 602/49\% & 1224\\
+\hline
+Chemistry & 125/8\% & 1370/92\% & 1495 & 55/3\% & 46/2\% & 1/0\% & 10/1\% & 12/1\% & 21/1\% & 295/16\% & 551/30\% & 866/47\% & 1857\\
+\hline
+Decision Sciences & 314/10\% & 2702/90\% & 3016 & 260/7\% & 21/1\% & 1/0\% & 20/1\% & 77/2\% & 28/1\% & 504/14\% & 988/27\% & 1766/48\% & 3665\\
+\hline
+Dentistry & 0/0\% & 19/100\% & 19 & 0/0\% & 0/0\% & 0/0\% & 1/4\% & 0/0\% & 1/4\% & 1/4\% & 11/44\% & 11/44\% & 25\\
+\hline
+Earth and Planetary Sciences & 481/23\% & 1587/77\% & 2068 & 159/6\% & 300/12\% & 22/1\% & 11/0\% & 41/2\% & 19/1\% & 289/11\% & 690/27\% & 1003/40\% & 2534\\
+\hline
+Economics, Econometrics and Finance & 97/45\% & 118/55\% & 215 & 70/27\% & 7/3\% & 0/0\% & 4/2\% & 23/9\% & 18/7\% & 14/5\% & 60/23\% & 68/26\% & 264\\
+\hline
+Energy & 54/4\% & 1330/96\% & 1384 & 26/2\% & 20/1\% & 1/0\% & 1/0\% & 6/0\% & 9/1\% & 124/7\% & 606/36\% & 883/53\% & 1676\\
+\hline
+Engineering & 394/3\% & 14323/97\% & 14717 & 198/1\% & 175/1\% & 5/0\% & 10/0\% & 43/0\% & 35/0\% & 2918/17\% & 5051/29\% & 9090/52\% & 17525\\
+\hline
+Environ Science & 790/45\% & 949/55\% & 1739 & 501/22\% & 77/3\% & 1/0\% & 69/3\% & 226/10\% & 249/11\% & 103/4\% & 480/21\% & 586/26\% & 2292\\
+\hline
+Health Professions & 94/14\% & 593/86\% & 687 & 62/7\% & 5/1\% & 1/0\% & 6/1\% & 21/2\% & 42/5\% & 143/17\% & 234/28\% & 331/39\% & 845\\
+\hline
+Immunology and Microbiology & 230/62\% & 141/38\% & 371 & 141/30\% & 28/6\% & 3/1\% & 11/2\% & 46/10\% & 73/15\% & 21/4\% & 75/16\% & 74/16\% & 472\\
+\hline
+Materials Science & 84/3\% & 3069/97\% & 3153 & 34/1\% & 47/1\% & 1/0\% & 0/0\% & 5/0\% & 5/0\% & 673/18\% & 1127/30\% & 1905/50\% & 3797\\
+\hline
+Mathematics & 753/9\% & 7878/91\% & 8631 & 607/6\% & 114/1\% & 6/0\% & 31/0\% & 156/2\% & 47/0\% & 1928/19\% & 2532/25\% & 4812/47\% & 10233\\
+\hline
+Medicine & 887/22\% & 3062/78\% & 3949 & 572/12\% & 88/2\% & 6/0\% & 71/1\% & 154/3\% & 286/6\% & 548/11\% & 1393/28\% & 1793/37\% & 4911\\
+\hline
+Multi-disciplinary & 330/36\% & 594/64\% & 924 & 181/16\% & 41/4\% & 3/0\% & 27/2\% & 76/7\% & 111/10\% & 93/8\% & 251/22\% & 378/33\% & 1161\\
+\hline
+Neuroscience & 563/38\% & 910/62\% & 1473 & 321/17\% & 70/4\% & 5/0\% & 35/2\% & 85/5\% & 239/13\% & 219/12\% & 326/18\% & 538/29\% & 1838\\
+\hline
+Nursing & 25/69\% & 11/31\% & 36 & 16/33\% & 0/0\% & 0/0\% & 3/6\% & 1/2\% & 15/31\% & 1/2\% & 5/10\% & 8/16\% & 49\\
+\hline
+Pharmacology, Toxicology and Pharmaceutics & 74/33\% & 148/67\% & 222 & 60/21\% & 5/2\% & 0/0\% & 4/1\% & 16/6\% & 8/3\% & 26/9\% & 67/24\% & 97/34\% & 283\\
+\hline
+Physics and Astronomy & 494/9\% & 4812/91\% & 5306 & 134/2\% & 324/5\% & 31/0\% & 8/0\% & 23/0\% & 31/0\% & 906/14\% & 1928/30\% & 3062/47\% & 6447\\
+\hline
+Psychology & 798/87\% & 119/13\% & 917 & 461/37\% & 32/3\% & 1/0\% & 62/5\% & 159/13\% & 397/32\% & 18/1\% & 51/4\% & 75/6\% & 1256\\
+\hline
+Social Sciences & 695/23\% & 2330/77\% & 3025 & 437/12\% & 44/1\% & 3/0\% & 48/1\% & 145/4\% & 295/8\% & 474/13\% & 806/22\% & 1474/40\% & 3726\\
+\hline
+Veterinary & 35/81\% & 8/19\% & 43 & 21/37\% & 0/0\% & 0/0\% & 5/9\% & 12/21\% & 7/12\% & 0/0\% & 6/11\% & 6/11\% & 57\\
+\hline
+\end{tabular}
+}
+\begin{tablenotes}
+\begin{fill} 
+\bigskip
+Table shows the basic subject split between Bayesian modeling and deep learning with total for the two categories, following are the individual package counts with a 'detail totals' as well for the individual packages.
+Note that the totals do not match because an article can match more than one package and as result gets contributes to the count of multiple packages.
+On average an article's journal is classified into two subject categories.
+\end{fill}
+\end{tablenotes}
+\end{table}
+
+\begin{center}
+\begin{table*}[t]%
+\centering
+\caption{Packages and Scopus Queries\label{tab2}}
+\begin{tabular}{|c c|} 
+ \hline
+ Package & Scopus Query  \\ [0.5ex] 
+ \hline
+Stan &	REF((gelman AND hoffman AND stan) OR mc-stan.org) \\
+ \hline
+brms &	REF(brms AND burkner) \\
+\hline
+PyStan &	REF(pystan) \\
+ \hline
+ RStanArm &	REF(rstanarm) \\
+  \hline
+RStan	& REF(rstan AND NOT mit) \\
+ \hline
+PyMC &	REF(PyMC3 OR (PyMC* AND fonnesbeck)) \\
+ \hline
+TensorFlow &	REF(tensorflow) \\
+ \hline
+PyTorch &	REF(pytorch) \\
+ \hline
+Keras &	REF(Keras) \\
+ \hline
+Bayesian &	REF(((gelman AND hoffman AND stan) OR mc-stan.org) OR (brms AND burkner) OR \\
+ & (pystan) OR (rstanarm) OR (rstan AND NOT mit) OR (PyMC3 OR (PyMC* AND fonnesbeck))) \\
+ \hline
+Deep Learning &	REF((tensorflow) OR (pytorch) OR (Keras)) \\
+ \hline
+\end{tabular}
+\begin{tablenotes}
+\begin{center}\bigskip Packages with corresponding Scopus queries as used for citation counting.
+\end{center}
+\end{tablenotes}
+\end{table*}
+\end{center}
+
+\begin{figure}
+    \centering
+    \caption{Relative counts with and without computer science journals for Baysian and deep learning packages.}
+    \label{fig:my_label}
+    \includegraphics{CompSciVsNonCompSci}
+\end{figure}
+
+\end{document}
